---
# Start of options to set
report_title: |
  Strait of Georgia Pacific Herring (*Clupea pallasii*) stock update in 2024
assess_yr: "2024"
report_year: "2025"
release_month: "January"
report_number: "XXX"
meeting_date: "September XX, 2024"
region: "Pacific Region"
phone: "(250) 555-5555"
email: "DFO.PacificCSA-CASPacifique.MPO@dfo-mpo.gc.ca"
csa_address: "3190 Hammond Bay Rd., Nanaimo, BC, V9T 6N7"
report_title_french: |
  Mise à jour du stock de hareng du Pacifique (*Clupea pallasii*)
  du détroit de Géorgie en 2024
inuktitut_citation: |
  Inuktitut citation here
# [Mandatory section; less than 150 words.]
context: |
  Pacific Herring (*Clupea pallasii*) abundance in British Columbia is assessed
  using a spatially integrated statistical catch at age model fit to
  spawn survey indices, fishery age composition data, and commercial catches.
  This stock assessment evaluates abundance relative to
  the limit reference point, the upper stock reference, and
  fisheries management objectives.
  [...]
  We recommend a management procedure for the
  Strait of Georgia stock assessment region.
output:
 csasdown::fsar_word
link-citations: false
bibliography: bib/refs.bib
# End of options to set
title: ""
knit: (function(input, ...) {
       csasdown::render_sar()
      })
---

```{r libraries, echo = FALSE, message = FALSE}
# Add packages here
library(knitr)
library(rmarkdown)
library(tidyverse)
library(rosettafish)
library(csasdown)
library(here)
library(patchwork)
library(scales)
library(gfutilities)
# library(gfiscamutils)
# library(herringutils)
```

```{r setup, echo = FALSE}
# Set knitr options
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "knitr-figs-docx/",
  fig.asp = 0.618,
  fig.width = 7,
  echo = FALSE,
  dev = "png",
  dpi = 180
)
```

```{r metadata}
# assess_yr <- metadata$assess_yr
assess_yr <- 2024
# meta_out <- metadata$output
# french <- meta_out$`csasdown::fsar_word`$french
french <- FALSE
```

```{r models}
models_dir <- here("models")
major_stock_dir <- list("SoG")
```

```{r scripts}
source(here("R", "models.R"))
source(here("R", "variables.R"))
source(here("R", "figures.R"))
source(here("R", "themes.R"))
# Theme for figures
theme_set(theme_herring())
```

<!-- Math expressions -->
\newcommand{\mli}[1]{\mathit{#1}}
\newcommand{\SB}{\mli{`r en2fr("SB", translate = french)`}}
\newcommand{\BR}{\mli{BR}}
\newcommand{\AVE}{\text{`r en2fr("AVE", translate = french)`}}

<!-- Compile the document on the command line: -->
<!-- bookdown::render_book("index.Rmd") -->

<!--chapter:end:index.Rmd-->

## SCIENCE ADVICE {-}

[Mandatory.
Maximum length 500 words.
Include summary bullets of the following factors affecting the stock
that are relevant for decision-making.
Please refer to the Guidance Document for standardized language and examples,
and what to avoid in this section.] 

### Status {-}

::: {custom-style="fsar-bullet"}
[The Strait of Georgia (SOG) major stock assessment region (SAR) is healthy.]
:::

define the word SAR 

### Trends {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)] 
:::

### Ecosystem and Climate Change Considerations {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)]
:::

### Stock Advice {-}

::: {custom-style="fsar-bullet"}
[Mandatory bullet(s)] 
:::

### Other Management Questions [if applicable, otherwise do not include] {-}

::: {custom-style="fsar-bullet"}
[Other bullets if applicable]
:::

<!--chapter:end:english/01_advice.Rmd-->

## BASIS FOR ASSESSMENT {-}

### Assessment Details {-}

<!-- [Mandatory] -->

Pacific Herring in BC are managed as
`r number_to_word(length(major_regions_short))` major and
`r number_to_word(length(minor_regions_short))` minor
SARs (Figure XX), and this report present science advice for the
Strait of Georgia (SoG) major SAR, and the other major and minor
SARs are presented in REF(2024SR).

SoG Pacific Herring are assessed using a spatially integrated statistical 
catch at age herring (SISCAH) modelling framework within a management 
strategy evaluation (MSE) process [@dfo2023a], fitted using
spawn survey indices, fishery age composition data, and commercial catches.

For SoG Herring, the SISCAH operating model includes: density dependent
mortality (DDM), integrated surface and dive survey data in the index,
commercial fishery timing, and a likelihood function with correlated 
age-compositions.

A weighted operating model ensemble is used for the management procedure 
evaluation, whereby the ensemble represents a range of productivity 
characterized by the stock recruitment steepness ($h$) and the lower 
limit on natural mortality ($M_b$). This approach better represents the 
uncertainty in stock productivity rather than using a single estimated
value of $h$.

Management procedures (MPs) were simulation tested using a 15-year projection 
period (2024-2038); only MPs that meet the established conservation 
objective [@dfo2022a] are presented.


#### Year Assessment Approach was Approved {-}

<!-- [Mandatory; year and reference] -->
2023 [@dfo2023a, @johnson2024].

#### Assessment Type {-}

<!-- [Mandatory. -->
<!-- Select one of the following: -->
Full Assessment and MSE process.
<!-- Interim Year Update, or -->
<!-- Harvest Decision Rule (HDR) Update. -->
<!-- See the Guidance Document for definitions.]  -->

#### Most Recent Assessment Date {-}

1.	Last Full Assessment: 2024 (this document).

<!-- 1.	Last Interim Year Update: [date and reference] -->

#### Assessment Approach {-}

1. Broad category:
<!-- [Select one of the following: -->
<!-- Data-poor (e.g., catch-MSY, very limited data, etc.), -->
<!-- index-based (trends in empirical indices only), -->
<!-- single stock assessment model, -->
<!-- model ensemble, -->
Single stock operating model ensemble (full MSE process).
<!-- or MSE-lite), -->
<!-- Wild Salmon Policy Integrated Assessment, -->
<!-- multiple approaches, -->
<!-- other (specify)] -->

1.	Specific category:
<!-- [Select one or as many as apply: -->
<!-- No data, -->
<!-- Index-based (fishery-dependent indices only), -->
index-based,
<!-- (including fishery-dependent and fishery-independent indices), -->
<!-- Delay difference, -->
<!-- Biomass dynamics, -->
<!-- Virtual population analysis, -->
<!-- Statistical catch-at-length, -->
statistical catch-at-age, 
density dependent time varying natural mortality, and
<!-- State-Space, -->
<!-- Sequential Population Analysis, -->
<!-- Surplus Production, -->
stock-recruitment relationship.
<!-- Potential Biological Removal, -->
<!-- Other (please explain)] -->

<!-- [Optional short summary of stock assessment methodology. -->
<!-- Include the method of considering ecosystem variables. -->
<!-- Include a reference to a Research Document or other publication, -->
<!-- and/or footnote to the working paper. -->
<!-- Maximum length 100 words.] -->

<!-- Mandatory -->
This `r assess_yr` SISCAH modelling framework [@johnson2024] includes stock 
status and MP for SoG Herring.

### Stock Structure Assumption {-}

<!--Stock overview information: -->
<!--[reference to a Research Document or other publication, and/or footnote to the working paper] -->
<!--[Mandatory. Maximum length 150 words.]-->

Stock overview information:
SoG Herring is one of five major stocks managed in British Columbia. 
Data collection and assessment focuses on the core spawn timing associated with 
the main migratory stock. Adult herring are reported in various areas of the Salish
Sea during the summer months indicating some level of non-migratory life history,
with the largest component of the stock migrating to the west coast to share summer/ fall
feeding grounds with the west coast Vancouver Island Herring stock.

### Reference Points {-}

A biological LRP was defined for Pacific Herring major stocks as 
$0.3\SB_0$ in 2018, where $\SB_0$ is estimated unfished spawning biomass [@kronlund2017]. 
Beginning in 2019, the LRP has been implemented 
in a conservation objective as follows:

Maintain spawning biomass at or above the LRP with at least 75% probability 
over three Pacific Herring generations
(i.e., avoid a biomass limit; $P(\SB_t \geq 0.3 \SB_0) \geq 0.75$)."

Candidate USRs were first introduced in @cleary2018 and a formal analysis of 
USR options was undertaken in 2022 with results presented in @dfo2022a.

In 2023, DFO Resource Management selected and implemented a provisional 
USR of $\overline{\SB}$ through 2022/23 IFMP process.

<!-- [Mandatory. -->
<!-- A table may be used in lieu of bullets for this section. -->
<!-- Refer to the Guidance Document for further instructions.] -->

- Limit Reference Point (LRP): $0.3\SB_0$ [@kronlund2017],
where $\SB_0$ is estimated unfished spawning biomass.

- Upper Stock Reference (USR): average spawning biomass
from `r yr_range[1]` to `r yr_range[2]` $\overline{\SB}$
(provisional; IFMP).

- Removal Reference (RR): not yet determined

- Target (TRP): not yet determined (USR=TRP)

### Stock Status {-}

OM status...
(Table X).

#### Other Stock Reference Points {-}

<!-- [If applicable (otherwise, do not include). -->
<!-- A table may be used in lieu of bullets for this section.]  -->

Currently, a core set of fisheries management objectives [@dfo2019c]
have been applied to each major SAR:

<!-- TODO: define one generation -->
1. Maintain spawning biomass at or above the LRP with
at least 75% probability over three Pacific Herring generations
(i.e., avoid a biomass limit; $P(\SB_t \geq 0.3 \SB_0) \geq 0.75$),

1. Maintain spawning biomass at or above the USR
with at least 50% probability over three Pacific Herring generations
(i.e., achieve a target biomass; $P(\SB_t \geq \SB_{targ}) \geq 0.50$),

1. Maintain average annual variability (AAV) in catch
below 25% over three Pacific Herring generations
(i.e., minimize catch variability; $\text{AAV} < 0.25$), and

1. Maximize average annual catch
over three Pacific Herring generations
(i.e., maximize average catch).

However, a fully specified set of objectives has not yet been developed
for each SAR.
DFO will continue to collaborate with coastal First Nations to
develop area-specific objectives specific to
FSC fisheries as well as SOK fisheries.
In addition, DFO will continue to engage with
the herring industry, government, and non-government organizations
to describe broader objectives related to conservation, economics, and access.

#### Habitat or Ecosystem Reference Points {-}

[If applicable (otherwise, do not include).
A table may be used in lieu of bullets for this section.] 

### Harvest Decision Rule {-}

[Optional (otherwise, do not include).
Maximum length 100 words.]

### Enhancement Plan {-}

[Optional (otherwise, do not include).
Maximum length 100 words.]

### Habitat Restoration Plan {-}

[Optional (otherwise, do not include).
Maximum length 100 words.]

### Data {-}

<!-- [Mandatory. -->
<!-- Maximum length 300 words.]  -->

<!-- Data changes: -->
<!-- [Optional (otherwise, do not include).] -->

There are three types of input data used for
the Pacific Herring stock assessment:
catch data, biological data, and abundance data.
<!-- These data are described in the following sections, and -->
<!-- summarized in Table XX. -->
Relative to the previous assessment,
the only change to input data was to extend all the time series
to include the `r this_season` herring season (July 1 to June 30).
Note that we refer to 'year' instead of 'herring season' in this report;
therefore `r assess_yr` refers to the `r this_season` Pacific Herring season.

<!--chapter:end:english/02_basis.Rmd-->

## ASSESSMENT {-}

- Example cross-reference:
Stock status in relation to four indicators (see Figure \@ref(fig:indicators)).

[Mandatory panel of images comes first – example shown.
Refer to the Guidance Document for options and further instructions.]

(ref:indicators-cap) Stock status indicators for Pacific Herring in the Strait of Georgia (SoG) major stock assessment region (SAR) from `r yr_range[1]` to `r yr_range[2]`. (A) Catch (line) and total allowable catch (TAC; dashed line). (B) Spawning stock biomass (SSB) in relation to the limit reference point (LRP; dashed red line) and upper stock reference (USR; dotted green line). (C) Fishing mortality (F; black line), removal reference (RR; dashed red line), and natural mortality (M; dotted line). (D) Recruitment of age-`r age_recruit` fish. Legend: lines indicate median values, grey shading indicates `r ci_pct`% confidence intervals, mortality is per year (yr), biomass is in tonnes (t), and recruitment is number of fish.

```{r indicators, fig.cap = "(ref:indicators-cap)"}
# Four-panel indicator plot
plot_indicators(dat = sim_dat)
```

[Optional. Add an additional optional panel of 4 indicators
if relevant to developing advice.] 
[Optional. If the FSAR or FSRR is being used
to provide advice for more than one stock,
the mandatory panel can be repeated for each stock.]

### Historical and Recent Stock Trajectory and Trends {-}

[Mandatory.
Maximum text length one page.
Must be first section in Assessment section.]

[Report trends in relevant indicator(s), one per paragraph,
in the following format:]

Insert `stock_status_text()` function to automate text:

- Estimated unfished spawning biomass $\SB_0$ is 138,491 t, and
the LRP of $0.3\SB_0$ is 41,547 t (posterior medians).
Compared to last year, estimated spawning biomass in 2023 $\SB_{2023}$
decreased from 75,523 to 72,782 t (posterior median), and
is equivalent to 52.1% of $\SB_0$ (Tables XX & XX).
Spawning biomass in 2023 is estimated to be
above the LRP with a 94.3% probability (Table XX).

#### [Indicator 1] [specify name, e.g., Abundance]: {-}

[Provide values and trends for the stock (as applicable).]

#### Unknown {-}

[Optional.
Indicate what indicators are not available for the stock (as applicable);
otherwise, do not include.]

### History of [Management; Landings/Harvest/Effort, TAC \& Catch Advice] {-}

[Mandatory.
Title can vary.
Maximum text length one page.] 

Harvest options for `r assess_yr+1` reflect application of
simulation-tested MPs for each major SAR,
derived from the Herring OM [@benson2022]. 
OM conditioning was updated in 2022 using
historic stock and fishery data from `r major_start_yr` to 2021;
no MP updates were conducted for 2023.
MPs are not provided for HG because this is now conducted within the 
HG rebuilding plan process.^[Haida Gwaii 'íináang | iinang Pacific Herring:
An ecosystem overview and ecosystem-based rebuilding plan.
Draft consultation period occurred September 26 to December 16, 2022; 
Final approval scheduled for fall 2023.]
<!-- TODO: Need to update rebuilding plan reference -->

<!-- #### History of Hatchery and Supplementation {-} -->

<!-- [If applicable (otherwise, do not include).] -->

<!-- #### History of Freshwater Habitat Impacts {-} -->

<!-- [If applicable (otherwise, do not include).] -->

### Projections [optional titles: Simulations, Future Scenarios] {-}

[Mandatory.
Title can vary.
Tables or figures can be used as necessary to support decision-making.]

Insert `proj_biomass_text()` function to automate text:

- In the absence of fishing,
spawning biomass in 2024 $\SB_{2024}$ is forecast
to be 73,375 t (posterior median; Table XX).
Spawning biomass in 2024 is forecast to be
below the LRP of $0.3\SB_0$ (41,547 t) with a 8.0% probability,
in the absence of fishing (Table XX and Figure XX).

### Ecosystem and Climate Change Considerations {-}

[Mandatory.
Maximum 300 words.
May be moved earlier in the Assessment section.]

### Harvest Decision Rule Outputs {-}

[If applicable (otherwise, do not include).] 

### Evaluation of Exceptional Circumstances/Assessment Triggers {-}

[If applicable (otherwise, do not include).] 

<!--chapter:end:english/03_assessment.Rmd-->

## BYCATCH {-}

<!-- In order to advance progress towards -->
<!-- a holistic ecosystem-based management (EBM) approach for Pacific Herring, -->
<!-- we present information to describe indirect effects -->
<!-- on herring populations such as incidental mortality.  -->
Some fisheries and aquaculture activities in BC cause
incidental mortality to Pacific Herring.
Similar to SOK harvest as well as FSC and recreational catch,
incidental mortality is currently only indirectly accounted for 
in Pacific Herring stock assessment by
estimating annual rates of natural mortality, 
and is considered minor relative to commercial harvest.
<!-- Impacts of additional mortality sources will be explored with the new  -->
<!-- modelling framework. -->

<!-- Again this year, we include data for -->
<!-- incidental mortality in finfish aquaculture activities -->

<!-- (Figure XX). -->
<!-- Note that incidental mortality data are reported by -->
<!-- Pacific Fishery Management Area (PFMA) -->
<!-- which are analogous to Statistical Areas, but -->
<!-- PFMAs are larger and can extend beyond SAR boundaries. -->
<!-- Data indicate the number of Pacific Herring dead and released, -->
<!-- with the following caveats: -->

<!-- 1. Unknown mortality rate for "released" fish, -->

<!-- 1. Unknown length, weight, age, and sex of fish, and -->

<!-- 1. "Herring-like" fish are assumed to be herring -->
<!-- when decomposition hinders identification. -->

<!-- Note that data from previous years are updated  -->
<!-- as DFO receives additional reports. -->

<!--chapter:end:english/04_bycatch.Rmd-->

## PROCEDURE FOR INTERIM YEAR UPDATES {-}

[Optional (otherwise, do not include).] 

<!--chapter:end:english/05_interim.Rmd-->

## OTHER MANAGEMENT QUESTIONS {-}

[Optional (otherwise, do not include).] 

<!--chapter:end:english/06_other.Rmd-->

## SOURCES OF UNCERTAINTY {-}

[Mandatory.
Maximum length 500 words.]

### Research Recommendations {-}

[Optional (otherwise, do not include).] 

<!--chapter:end:english/07_uncertainty.Rmd-->

## LIST OF MEETING PARTICIPANTS {-}

[Mandatory.
The list is to include the name of each participant and their affiliation.]

<!--chapter:end:english/08_participants.Rmd-->

## SOURCES OF INFORMATION {-}

<!-- [Mandatory.] -->

<!-- [References should only be included to understand the advice, and -->
<!-- should not be included to give credit to the work -->
<!-- as they are done in the Research Document. -->
<!-- References to the unpublished working paper should be -->
<!-- integrated as footnotes through this document.] -->

<!-- [References must be formatted according to -->
<!-- the Guide for the Production of Fisheries and Oceans Canada Reports.] -->

<!-- [References listed in the References Cited section should be -->
<!-- publicly available when the report is posted on the website. -->
<!-- The only exception is for books or articles that -->
<!-- have been accepted for publication -->
<!-- (page proof stage or where there is an acceptance letter). -->
<!-- These can be listed followed by the notation “in press”. -->
<!-- All others should be included in the text -->
<!-- as unpublished data or working papers.] -->

<!--chapter:end:english/09_sources.Rmd-->

