## BASIS FOR ASSESSMENT {-}

### Assessment Details {-}

The Pacific Herring Strait of Georgia 
(SoG) stock assessment region, is one of the 
`r number_to_word(length(major_regions_short))` major stock areas; the other 
major and the `r number_to_word(length(minor_regions_short))` minor stock
areas are presented in [@dfo2024bInPrep, In Prep.].

Managment strategies for SoG Pacific Herring are evaluated with a management 
strategy evaluation (MSE) process, which implements a spatially integrated statistical 
catch at age Herring (SISCAH) modelling framework [@dfo2023a]. 
The modelling framework is fitted to historical time series of spawn survey indices, 
biological and fishery age composition data, and commercial fishery catches.

The SISCAH operating model estimates SoG Herring historical biomass time series 
from 1951-2003 with these features: density dependent mortality (DDM), an integrated surface 
and dive survey index, gear-specific commercial fishery timing, and a 
likelihood function with correlated age-compositions.
Incorporation of density dependent natural mortality models potential ecosystem 
impacts (e.g., depensatory predation) on Pacific Herring stocks in line with
DFO's EAFM (ecosystem approaches to fisheries management).

A weighted ensemble of five differently-parameterized SISCAH operating models 
is used for management procedure evaluation. 
This approach was chosen to encompass the uncertainty around the 
productivity parameters stock recruitment steepness 
($h$), which affects productivity at low stock sizes, and the lower 
limit on natural mortality ($M_b$), which represents the average mortality rate 
at high stock sizes. Uncertainty around $h$ is especially high since abundance in
SoG has rarely dipped low enough to give the model a strong indication of what 
$h$ should be. Details of the operating model ensemble are described in 
Appendix [A](#sec:app-a).

The weighted ensemble was used to simulate management procedures (MPs) over a 
15-year projection period (2024-2038) and present a MP tuned to meet 
the conservation objective
(Section "[Fishery management objectives](#sec:fish-mgmt-objs)") that is
compliant with both the “DFO Sustainable Fisheries Framework” policy and 
“A fishery decision-making framework incorporating the Precautionary Approach”
policy as per the 
[Terms of Reference](https://www.dfo-mpo.gc.ca/csas-sccs/Schedule-Horraire/2024/09_23-eng.html).

#### Year Assessment Approach was Approved {-}

The SISCAH operating model framework implemented here for SoG Herring was 
reviewed and approved by CSAS in June 2023 [@dfo2023a].

#### Assessment Type {-}

<!-- [Mandatory. -->
<!-- Select one of the following: -->
Full assessment,
MSE process,
with update to the management procedure and harvest control rule.
<!-- Interim Year Update, or -->
<!-- Harvest Decision Rule (HDR) Update. -->
<!-- See the Guidance Document for definitions.]  -->

#### Most Recent Assessment Date {-}

Last Full Assessment: this document (2024).

#### Assessment Approach {-}

1. Broad category:
Full MSE process using a single stock operating model ensemble.
<!-- [Select one of the following: -->
<!-- Data-poor (e.g., catch-MSY, very limited data, etc.), -->
<!-- index-based (trends in empirical indices only), -->
<!-- single stock assessment model, -->
<!-- model ensemble, -->
<!-- or MSE-lite), -->
<!-- Wild Salmon Policy Integrated Assessment, -->
<!-- multiple approaches, -->
<!-- other (specify)] -->

2. Specific category:
Index-based,
state-space,
statistical catch-at-age model,
with a Beverton-Holt stock-recruitment relationship and
density dependent time varying natural mortality.
<!-- [Select one or as many as apply: -->
<!-- No data, -->
<!-- Index-based (fishery-dependent indices only), -->
<!-- (including fishery-dependent and fishery-independent indices), -->
<!-- Delay difference, -->
<!-- Biomass dynamics, -->
<!-- Virtual population analysis, -->
<!-- Statistical catch-at-length, -->
<!-- Sequential Population Analysis, -->
<!-- Surplus Production, -->
<!-- Potential Biological Removal, -->
<!-- Other (please explain)] -->

<!-- [Optional short summary of stock assessment methodology. -->
<!-- Include the method of considering ecosystem variables. -->
<!-- Include a reference to a Research Document or other publication, -->
<!-- and/or footnote to the working paper. -->
<!-- Maximum length 100 words.] -->

<!-- Mandatory -->
This SISCAH modelling framework [@johnson2024] estimates SoG Herring stock status 
and quota calculations for 2024/25 season using the MP tuned to the conservation 
objective.

### Stock Structure Assumption {-}

<!--Stock overview information: -->
<!--[reference to a Research Document or other publication, and/or footnote to the working paper] -->
<!--[Mandatory. Maximum length 150 words.]-->

SoG Herring is one of five major BC Pacific Herring stocks [@dfo2024].
The SoG stock spawns primarily in March, over multiple locations with variable timing.
Details of spawn timing and locations are described in the
[2024 SoG data summary report](https://github.com/grinnellm/Reports/blob/master/Summaries/2024/DataSummary.SoG.2024.pdf).
Post spawning, mature SoG Herring
migrate to the west coast Vancouver Island (WCVI) where they share summer/ fall 
feeding grounds with herring that spawn on the WCVI. 
However adult herring are also reported in the 
Salish Sea throughout the summer months indicating some level of non-migratory life
history. Previous genetic research has not identified finer-scale stock structure 
within the SoG management area, however, new genetic methods are currently being 
used to update baseline data for SoG spawning ares and this may provide new 
insights.

### Reference Points {-}

<!-- [Mandatory. -->
<!-- A table may be used in lieu of bullets for this section. -->
<!-- Refer to the Guidance Document for further instructions.] -->
A suite of biological reference points are used to develop performance metrics 
for the MP evaluations. 
These include a range of proportions of $B_0$ (unfished spawning biomass) and 
$B_{MSY}$ (biomass at maximum sustainable yield). 
Figure \@ref(fig:yieldSSBcurve) shows the relative position of biological reference
points for SISCAH operating model 1 (OM 1, Appendix [A](#sec:app-a)),

```{r yieldSSBcurve, echo = FALSE, fig.cap = yieldSSBcap}
yieldSSBcap <- "Strait of Georgia Herring operating model 1 (OM 1) equilibrium yield curve with respect to spawning stock biomass. The horizontal dashed line indicates OM1's estimate of maximum sustainable yield ($MSY$), while vertical line segments show biomass reference points including, from right to left, unfished biomass ($B_0$), the provisional upper stock reference (USR= 0.8$B_{1998:2007}$), optimal biomass producing $MSY$ on average ($B_{MSY}$), default USR from the DFO policy (80\\% of $B_{MSY}$), the Herring limit reference point (LRP= 0.3$B_0$), and the default LRP from the DFO policy ($0.4 B_{MSY}$)."
filepath <- "data/yieldSSB.png"
knitr::include_graphics(filepath)
```

#### Limit Reference Point (LRP) {-}

The LRP is $0.3B_0$, = `r round(ensRefPtsTable$B0*0.3, 1)` kt where ${B_0}$ is the long-term
average estimated unfished spawning biomass [@kronlund2017],[@forrest2023].
This LRP is more conservative than the default LRP identified in 
“DFO Sustainable Fisheries Framework” policy [@dfo2009].

#### Upper Stock Reference (USR) {-}

The provisional USR is `r round(ensRefPtsTable$Busr, 1)` kt.
Candidate USRs were first introduced in @cleary2018 and a formal analysis of 
USR options was undertaken in 2022 [@dfo2022a].
Following consultations, DFO Resource Management then selected and implemented 
0.8$B_{1998:2007}$ as a "provisional USR" for the 2022/23 IFMP 
where the USR was included as a performance metric in the MP evaluations.
Because consultations are ongoing for SoG Herring USRs, we include 
5 USR performance metrics in this analysis: 0.8$B_{MSY}$ (new this year) and 
all four candidate USRs presented as $B_{MSY}$ proxies [@dfo2022a].

#### Target Reference Point (TRP) {-}

A biomass target objective centered on the USR has been suggested within the 
Herring MSE process, however final selection of a USR for SoG Herring requires 
additional consultations and thus a TRP has not been defined.

#### Removal Reference (RR) {-}

The default maximum RR for commercially harvested fish stocks is
$F_{MSY}$ [@dfo2009, @dfo2013].
We currently do not specify an additional RR for SoG Herring because we 
use a process that directly selects an MP that meets the 
conservation objective, avoids the LRP with high probability, 
and whose harvest rate is more conservative than $U_{MSY}$).

#### Ecosystem Reference Points {-}

Ecosystem reference points are incorporated into
the conservation objective for Herring through:

1. A LRP of 0.3${B_0}$ based on an analysis of stock productivity [@kronlund2017]
which is well-above the default LRP of 0.4$B_{MSY}$ [@dfo2009] and incidentally 
is also above the more conservative soft-limit 
of 0.5$B_{MSY}$ implemented in New Zealand fisheries [@shelton2008].
The LRP of 0.3${B_0}$ for the ensemble operating model is
approximately 0.7$B_{MSY}$ (Appendix [A](#sec:app-a)).

2. Incorporating density dependent mortality to account for predation mortality,
representing the trophic level of Herring as a forage species.

### Fishery Management Objectives {-#sec:fish-mgmt-objs}

The first and core of the fisheries management objectives [@dfo2019c] is 
centred on the LRP and labelled the conservation objective:

1. Maintain spawning biomass at or above the LRP with
at least 75% probability over three Pacific Herring generations
(i.e., avoid a biomass limit; $P(\SB_t \geq 0.3 \SB_0) \geq 0.75$).

Three additional objectives are presented for MP evaluation only if the 
LRP requirement is met:

2. Maintain spawning biomass at or above the USR with at least 50% 
probability over three Pacific Herring generations (i.e., achieve a 
target biomass; $P(B_t \geq B_{targ}) \geq 0.50$),

3. Maintain average annual variability (AAV) in catch below 25% over three
Pacific Herring generations (i.e., minimize catch variability; 
$\text{AAV} < 0.25$), and

4. Maximize average annual catch over three Pacific Herring 
generations (i.e., maximize average catch).

DFO collaborates with coastal First Nations to develop 
area-specific objectives for Indigenous fisheries and 
engage with the Herring industry, government, and non-government 
organizations to describe broader objectives related to conservation, 
economics, and access.

### Operating Model Stock Status {-}

Stock status for the SoG major stock is derived from the operating model 
ensemble which represents weighting across five different models each 
representing a range of productivity levels (Appendix [A](#sec:app-a)).
The ensemble model stock status for 2023 is reported in Section 
"[Assessment under MSE](#sec:assessment)"

DFO Science, Pacific Region, will be updating operating models for SoG Herring 
every three years, following recommendations of @dfo2023a.
In 2026 SISCAH will be fit to new SoG survey and fishery data 
(extending the historical time series) and MPs will be re-evaluated against 
fishery management objectives. 
New objectives and performance metrics may
also be added to the MSE process at this time.

In the interim years (2024 and 2025), operating model status remains constant, 
and the MP is updated which estimates spawning biomass and the one-year 
projection required to calculate a TAC.

### Harvest Decision Rule {-}

Herring MSE presents sustainable harvest advice for SoG Herring using 
simulation tested MPs which identify the maximum target harvest rate for 
a MP tuned to the established conservation objective. 
MPs are evaluated using a 15-year projection period and 
simulated performance is reported against a suite of biological reference points
and fishery metrics.
See Section "[Projections](#sec:projections)" for a full description.

The harvest control rule component of each simulation-test MP is
a "hockey-stick" shaped function defined by lower and upper control points, and
a maximum target harvest rate [@cox2013roles]. 
Three shapes which differ in their 
control points and maximum target harvest rates
(Figure \@ref(fig:plotHCRules)) are presented for illustration purposes.
These options were chosen because previous results showed the harvest rate as
the key factor governing conservation performance for Herring. A ramp is 
implemented as per @dfo2009. 
Sustainable harvest advice for SoG Herring for 2024/25 is presented using 
the MP tuned to the conservation objective which has lower and upper control 
points at 30\% and 60\% of unfished spawning biomass and a maximum target
harvest ($maxTHR$) of 14\% (green solid line, Figure \@ref(fig:plotHCRules)).

```{r plotHCRules, echo = FALSE, fig.cap = HCRcap, fig.asp = 0.9}
HCRcap <- "Three harvest control rule components of SoG management procedures (MPs). The green line represents the MP tuned to meet the conservation objective over the 15-year projection period for the ensemble operating model (with density dependent natural mortality) and maximum target harvest rate, $maxTHR$, of 14\\%. The red line denotes an additional hockey-stick shaped rule with the previously used 20\\% maximum target harvest rate and the purple line shows the DFO default harvest control rule with maximum target harvest rate set at $U_{MSY}$ and control points at $0.4 \\cdot B_{MSY}$ and $0.8 \\cdot B_{MSY}$."

plotHCRules(  LCP_i = c(.3,.3,.172),
              UCP_i = c(.6,.6,.342),
              lowF = .0,
              highF = 1,
              U_i = c(.14,.20,.187) ,
              language = "English")
```

### Performance Evaluation for Interim Years and Exceptional Circumstances {-}

For each interim year the realized catch, survey index, and 
estimation-model-derived spawning biomass, are 
compared to the simulation envelopes from the ensemble model.
Exceptional circumstances are determined 
by visually comparing these three values for the new year
with the central 95\% of the ensemble operating model simulations.
Any unusual deviations from the ensemble model projections would be 
investigated and consideration given to refitting operating models.
For example, if the previous year's realized catch far exceeds the TAC 
prescribed by the MP or if the survey index is far below the projected index.

### Data {-}

<!-- [Mandatory. -->
<!-- Maximum length 300 words.]  -->
<!-- Data changes: -->
<!-- [Optional (otherwise, do not include).] -->

There are three types of input data used for
the Pacific Herring stock assessment:
catch data, biological data, and a fishery independent survey index.
The survey index is made up of two classes of surveys, surface
and dive. Combining the surveys into a blended index properly accounts for 
years in which these survey types co-occur and is a feature 
of the new SISCAH model.
SoG data are described in full in the [2024 SoG data summary report](https://github.com/grinnellm/Reports/blob/master/Summaries/2024/DataSummary.SoG.2024.pdf).
