## ASSESSMENT UNDER MSE {-#sec:assessment}

```{r calcTACchunk, echo = FALSE, include = FALSE}
TAC.df <- calcTAC(fit_maxTHR0.14, assessyear = assess_yr)

```

Stock trends and status for SoG Herring historical time series (1951-2023)
is represented by the ensemble operating model using 6 indicators (Figure \@ref(fig:indicators)).
Ensemble operating model trends are similar to those presented for SoG Herring 
in previous stock assessments (e.g., [@dfo2024]), with  
near-identical data sources and choice of indicators.
Step wise comparisons are found in [@johnson2024].

<!--TODO (Sam): 
For ref:indicators-cap:
1. the autodates need to state 1951-2023. I'm not sure where to change `r yr_range[1]` to `r yr_range[2]` 

SDNJ: Done. Good catch - this was taking years off the dummy table of simulated data

2. please confirm age-`r age_recruit` fish is age-1 (I thought this was age-2 but maybe I'm wrong)

SDNJ: It's age-1, ISCAM used to estimate age-2, but SISCAH estimates age-1.

3. please add average recruitment to (E)

SDNJ: Done

4. (F) Can you reverse the heat map to show darkest colour as most recent? 

SDNJ: Done.
-->


(ref:indicators-cap) Estimated stock status indicators for Pacific Herring in the Strait of Georgia (SoG) major stock assessment region (SAR) from `r yr_range[1]` to `r yr_range[2]`, as estimated by the ensemble operating model. All lines indicate weighted ensemble posterior median values, and where applicable, shading indicates `r ci_pct`\% credibility intervals. (A) Catch (black solid line). (B) Estimated spawning stock biomass (x 1,000 t) with limit reference point (LRP; dashed red line), productive period upper stock reference (USR; dashed green line), and unfished spawning biomass ($B_0$; dashed black line). (C) Estimated natural mortality rate ($M$ /yr; red line). (D) Estimated harvest rate ($U_t$, black trend line) with dashed horizontal line at 20\% shown from 1983-2022). (E) Recruitment of age-`r age_recruit` fish, with the historical average shown as a black dashed line. (F) Estimated surplus production (vertical axis) vs. spawning biomass (horizontal axis) with lighter points representing earlier years. vertical dashed lines denote the LRP (red), productive period USR (green), and $B_0$ (black).

```{r indicators, fig.cap = "(ref:indicators-cap)", fig.asp = 1.1}
# Six-panel indicator plot
baseplot_indicators(dat = true_dat)
```

### Historical and Recent Stock Trajectory and Trends {-}

Estimated unfished spawning biomass $\SB_0$ is 138,491 t, and
the LRP of $0.3\SB_0$ is 41,547 t (posterior medians).
Compared to last year, estimated spawning biomass in 2023 $\SB_{2023}$
decreased from 75,523 to 72,782 t (posterior median), and
is equivalent to 52.1% of $\SB_0$ (Tables XX & XX).
Spawning biomass in 2023 is estimated to be
above the LRP with a 94.3% probability (Table XX). -->

```{r OM_stock_status, echo = FALSE}
# 
stock_status_text(refPtsTab = ensRefPtsTable,
                                parTab = ensParTable,
                                fYear = 1951, lYear = 2023 )
```

#### Spawning Biomass and Status Relative to Reference Points {-}
The ensemble model estimates spawning biomass to be above $B_{MSY}$ and
the USR for the majority of the historical time series 
(Figure \@ref(fig:indicators)A).
Additionally, median posterior estimates of spawning biomass have remained
well-above the LRP in all years since 1970.

#### Recruitment and Natural Mortality {-}
Most of the fluctuations in estimated spawning biomass have been attributed 
to estimated recruitment, with lower than average age-1 recruitment in the 1970s, 
1980s, and in 2007 and 2009, all of which mirror dips in spawning biomass. 
Above average recruitment occurred in the intervening times corresponding to
biomass peaks (Figure \@ref(fig:indicators)E). 
While 2020 is one of the three highest peaks in estimated spawning biomass 
(since 1970), rising estimates of natural mortality (Figure \@ref(fig:indicators)C)
since 2015 has moderated impacts of above average recruitment.
Finally, opposing fluctuations in estimated recruitment and natural mortality 
likely contribute to increased uncertainty in spawning biomass and 
forecast biomass (#FIG).

#### Spawn biomass production {-}

Estimated spawning biomass production was high in 2019 and then declined each 
of the next three year to 2023 (Figure \@ref(fig:indicators)F), despite spawning
biomass remaining fairly constant.

<!-- #### Fishing Mortality {-} 
JC: not including because not talking about RR-->

### Ecosystem and Climate Change Considerations {-}

Density dependence, where increased mortality is associated with smaller biomass
models ecosystem effects such as the increased percentage of natural mortality 
in the event of lower biomass, consistent with predator consumption. 

### History of [Management; Landings/Harvest/Effort, TAC \& Catch Advice] {-}

STOPPED HERE

[Mandatory. make this short & ] 

COVER:
-20% hr IMPLEMENTED SINCE 1983
-POINT READER TO THE 6 PANEL FIGURE
-MENTION SIMULATION TESTING OF MPs FOR SoG HAS BEEN ONGOING 
SINCE 2015 - cite past herringSR


### Projections and MP Outputs for 2025 {-#sec:projections}

-JC: explain the process under new MSE
REMIND READER THAT THE ENSEMBLE OM REPRESENTS THE HISTORY AS 1951-2023, AND THEN THAT 2024, 2025,
AND 2026 ARE INTERIM YEARS IN WHICH WE DO NOT REFIT THE OM BUT RATHER USE THE
ESIMTATION MODEL COMPONENT OF THE MP TO ESTIMATE STOCK STATUS (2024), 
PROJECT THE STOCK FORWARD (TO 2025), AND THEN FEED THAT ESTMATE TO THE HCR.

#### Interim Year Update for 2024 {-}
<!-- TODO (Sam): please create current year statement function using the estimation model for 2024.
"Estimated spawning biomass $SB_{2024}$ is 90.23 kt in 2024, and is equivalent to 87.2 \\% of $SB_0$ (Tables XX & XX). Spawning biomass in 2024 is estimated to be above the LRP with a 99 \\% probability (Table XX)."

```{r, current_year_statement, echo = FALSE, warnings = FALSE}
# current_biomass_text() 
```

#### MP {-}

HERRING MPs ARE COMPROSED OF THREE COMPONENTS: DATA, ESTIMATION MODEL, HARVEST CONTROL RULE

MPS THAT WE CONSIDER DIFFER ONLY BY SHAPE OF HCR (point to Figure \@ref(fig:plotHCRules) )

We simulation tested three MPs with hockey-stick shaped harvest control rules (#refSecThisDoc)
and present performance indicators for the MP tuned to meet the conservation objective and
that is compliant with both the “DFO Sustainable Fisheries Framework” policy and 
“A fishery decision-making framework incorporating the Precautionary Approach” policy
(as per the Terms of Reference (#refLinkCSASTOR)).

WE USE THE CONSERVATION OBJECTIVE AS THE MIN THRESHOLD FOR ACCEPTABLE MP...


<!-- TODO: define three generations == 15-years -->


#### Interim Year Updates {-}

<!-- TODO (Sam): pls show us where to find these variable names so we can add auto-text

SDNJ: For below variables:
tIdx = 64 for 2024
tIdx = 65 for 2025

You could also use

thisYrIdx <- assess_yr - 1951 + 1
nextYrIdx <- thisYrIdx + 1

B_2024 = fit_maxTHR0.14$repOpt$SB_pt[1,64]
B_2025 = fit_maxTHR0.14$repOpt$SB_pt[1,65]

B0 = fit_maxTHR0.14$repOpt$B0_p[1]

If you look want to find more stuff, follow the calcTAC() function in 
R/functions.R

I also already created a proj_biomass_text() function that did this,
you can edit it in R/functions.

JC: Sam to show us this.

-->

-JC: STATE THE EM STATUS FOR 2024
-Bt_2024 
-Bt_2024/Bo
-Bt_2025
-Bt_2025/Bo

```{r, forecast_statement, echo = FALSE, warnings = FALSE}
# TODO: (SAM) Please make this look pretty.
proj_biomass_text(fit_maxTHR0.14) 
```

### Harvest Decision Rule Outputs {-}

```{r, TACtable, echo = FALSE, warnings = FALSE}

TACtabCap <- "Harvest decision rule outputs when the SoG Herring SAR MP is applied to data up to 2024."

kable( TAC.df, escape = FALSE, 
        caption = TACtabCap, booktabs = TRUE,
        align = c("l","c"),
        linesep = "\\addlinespace")


```

```{r, MPcaptions, include = FALSE, echo = FALSE}

stockLabs <- fit_maxTHR0.14$stock
species   <- fit_maxTHR0.14$ctlList$ctrl$speciesName
stock     <- fit_maxTHR0.14$ctlList$ctrl$stockNames

SBtRtMtFtCaption <- paste("Time series of spawning biomass with scaled spawn 
indices (top),recruitments (second row), natural mortality (third row), and
harvest rates (bottom row) for substocks of ", species, ". Stocks are, from 
left to right,", paste(stockLabs, collapse = ", "),".", sep = "")


hcrCaption <- paste0(stock," Herring management procedure harvest control rule 
(line) and TAC advice (annotation).")


```

```{r AtAGlance, include = include_extras, echo = FALSE, message = FALSE, fig.cap = SBtRtMtFtCaption , fig.asp = 1.1}
plotMulti(ts = c("SBtIdx","Rt","Mt","Ftg"), repList = fit_maxTHR0.14)
```

```{r mpHCR, echo = FALSE, message = FALSE, fig.cap = hcrCaption, fig.asp = 1.1}
plotRule(repList = fit_maxTHR0.14)
```

### Evaluation of Exceptional Circumstances/Assessment Triggers {-}

Fishery monitoring data collected since the adoption of the current SoG
Herring MP indicate that the population is behaving within operating 
model expectations. In formal language there is no indication of operating
model mis-specification (Figure \@ref(fig:ECfig)), also know as exceptional
circumstances. Mis-specification is checked visually by comparing catch and 
spawn index values within the central 95% of 2023 ensemble operating 
model simulations.

<!-- TODO (Sam): 
For ECfig:
1. top - Is the grey envelope (post 2023) spawning biomass and the blue spawn index? Why is the
uncertainty in the index larger than SB?

SDNJ: Because the noise in the index is the uncertainty in SSB (process error) plus 
uncertainty in spawn index (observation error)

-->

```{r ECfig, echo = FALSE, fig.cap = ECfigCap, fig.asp = 1.1 }
ECfigCap <- "Evaluation of exceptional circumstances and operating model mis-specification: graphical comparison of simulation envelopes from the ensemble model and catch and spawn index data from interim year(s) f. Panels show the central 95\\% of projected simulated data, including spawn index data (top) and catch (bottom) under the weighted ensemble 2023 SoG operating model. Forward projections in time (15-years) represent the maxTHR0.14 MP. Realised data are overlaid as points for the history (black) and projection years (red)."

SOGdata <- read.csv("data/SOGIdxCatData.csv")

plotSimDataEnvelopes( obj = mpBlobList[[2]],
                      sIdx = 1, pIdx =1, fIdx = 5,
                      Cdata = SOGdata$Catch,
                      Idata = SOGdata$totSpawn,
                      fYear = 1951 )

```
