## ASSESSMENT {-}

```{r calcTACchunk, echo = FALSE, include = FALSE}
TAC.df <- calcTAC(fit_maxTHR0.14)

```


- Example cross-reference:
Stock status in relation to four indicators (see Figure \@ref(fig:indicators)).

[Mandatory panel of images comes first – example shown.
Refer to the Guidance Document for options and further instructions.]

(ref:indicators-cap) Stock status indicators for Pacific Herring in the Strait of Georgia (SoG) major stock assessment region (SAR) from `r yr_range[1]` to `r yr_range[2]`. (A) Catch (line) and total allowable catch (TAC; dashed line). (B) Spawning stock biomass (SSB) in relation to the limit reference point (LRP; dashed red line) and upper stock reference (USR; dotted green line). (C) Fishing mortality (F; black line), removal reference (RR; dashed red line), and natural mortality (M; dotted line). (D) Recruitment of age-`r age_recruit` fish. Legend: lines indicate median values, grey shading indicates `r ci_pct`\% credibility intervals, mortality is per year (yr), biomass is in tonnes (t), and recruitment is number of fish.

```{r indicators, fig.cap = "(ref:indicators-cap)"}
# Four-panel indicator plot
plot_indicators(dat = true_dat)
```

[Optional. Add an additional optional panel of 4 indicators
if relevant to developing advice.] 
[Optional. If the FSAR or FSRR is being used
to provide advice for more than one stock,
the mandatory panel can be repeated for each stock.]

### Historical and Recent Stock Trajectory and Trends {-}

[Mandatory.
Maximum text length one page.
Must be first section in Assessment section.]

[Report trends in relevant indicator(s), one per paragraph,
in the following format:]

`r stock_status_text()` 

<!-- function to automate text:

- Estimated unfished spawning biomass $\SB_0$ is 138,491 t, and
the LRP of $0.3\SB_0$ is 41,547 t (posterior medians).
Compared to last year, estimated spawning biomass in 2023 $\SB_{2023}$
decreased from 75,523 to 72,782 t (posterior median), and
is equivalent to 52.1% of $\SB_0$ (Tables XX & XX).
Spawning biomass in 2023 is estimated to be
above the LRP with a 94.3% probability (Table XX). -->

#### [Indicator 1] [specify name, e.g., Abundance]: {-}

[Provide values and trends for the stock (as applicable).]

#### Unknown {-}

[Optional.
Indicate what indicators are not available for the stock (as applicable);
otherwise, do not include.]

### History of [Management; Landings/Harvest/Effort, TAC \& Catch Advice] {-}

[Mandatory.
Title can vary.
Maximum text length one page.] 

Harvest options for `r assess_yr+1` reflect application of
simulation-tested MPs for each major SAR,
derived from the Herring OM [@benson2022]. 
OM conditioning was updated in 2022 using
historic stock and fishery data from `r major_start_yr` to 2021;
no MP updates were conducted for 2023.
MPs are not provided for HG because this is now conducted within the 
HG rebuilding plan process.^[Haida Gwaii 'íináang | iinang Pacific Herring:
An ecosystem overview and ecosystem-based rebuilding plan.
Draft consultation period occurred September 26 to December 16, 2022; 
Final approval scheduled for fall 2023.]
<!-- TODO: Need to update rebuilding plan reference -->

<!-- #### History of Hatchery and Supplementation {-} -->

<!-- [If applicable (otherwise, do not include).] -->

<!-- #### History of Freshwater Habitat Impacts {-} -->

<!-- [If applicable (otherwise, do not include).] -->

### Projections [optional titles: Simulations, Future Scenarios] {-}

[Mandatory.
Title can vary.
Tables or figures can be used as necessary to support decision-making.]

`r proj_biomass_text()`

<!-- Insert `proj_biomass_text()` function to automate text:

- In the absence of fishing,
spawning biomass in 2024 $\SB_{2024}$ is forecast
to be 73,375 t (posterior median; Table XX).
Spawning biomass in 2024 is forecast to be
below the LRP of $0.3\SB_0$ (41,547 t) with a 8.0% probability,
in the absence of fishing (Table XX and Figure XX).
 -->
<!-- TODO: define three generations == 15-years -->


### Ecosystem and Climate Change Considerations {-}

Density dependence, where increased mortality is associated with smaller biomass
models ecosystem effects such as the increased percentage of natural mortality 
in the event of lower biomass, consistent with predator consumption. 

### Harvest Decision Rule Outputs {-}



```{r, TACtable, include = FALSE, echo = FALSE}

TACtabCap <- "Harvest decision rule outputs when the SoG herring SAR MP is applied to data up to 2024."

kable( TAC.df, escape = FALSE, 
        caption = TACtabCap, booktabs = TRUE,
        align = c("l","c"),
        linesep = "\\addlinespace")


```



```{r, MPcaptions, include = FALSE, echo = FALSE}

stockLabs <- fit_maxTHR0.14$stock
species   <- fit_maxTHR0.14$ctlList$ctrl$speciesName
stock     <- fit_maxTHR0.14$ctlList$ctrl$stockNames

SBtRtMtFtCaption <- paste("Time series of spawning biomass with scaled spawn 
indices (top),recruitments (second row), natural mortality (third row), and
harvest rates (bottom row) for substocks of ", species, ". Stocks are, from 
left to right,", paste(stockLabs, collapse = ", "),".", sep = "")


hcrCaption <- paste0(stock," Herring management procedure harvest control rule 
(line) and TAC advice (annotation).")


```




```{r, AtAGlance, echo = FALSE, message = FALSE, fig.cap = SBtRtMtFtCaption , fig.asp = 1.1}
plotMulti(ts = c("SBtIdx","Rt","Mt","Ftg"), repList = fit_maxTHR0.14)
```

```{r mpHCR, echo = FALSE, message = FALSE, fig.cap = hcrCaption, fig.asp = 1.1}
plotRule(repList = fit_maxTHR0.14)
```



### Evaluation of Exceptional Circumstances/Assessment Triggers {-}

Fishery monitoring data collected since the adoption of the current SOG
herring MP indicate that the population is behaving within operating 
model expectations. In formal language there is no indication of operating
model mis-specification (Figure \@ref(fig:ECfig)), also know as exceptional
circumstances. Mis-specification is checked visually by comparing catch and 
spawn index values within the central 95% of 2023 ensemble operating 
model simulations.


```{r ECfig, echo = FALSE, fig.cap = ECfigCap, fig.asp = 1.1 }
ECfigCap <- "A graphical test for exceptional circumstances indicating operating model mis-specification. Panels show the central 95\\% of projected simulated data, including spawn index data (top) and catch (bottom) under the weighted ensemble 2023 SOG operating model, projected forward in time under the maxTHR0.14 MP. Realised data are overlaid as points for the history (black) and projection years (red)."

SOGdata <- read.csv("data/SOGIdxCatData.csv")

plotSimDataEnvelopes( obj = mpBlobList[[2]],
                      sIdx = 1, pIdx =1, fIdx = 5,
                      Cdata = SOGdata$Catch,
                      Idata = SOGdata$totSpawn,
                      fYear = 1951 )

```






