## ASSESSMENT UNDER MSE {-#sec:assessment}

```{r calcTACchunk, echo = FALSE, include = FALSE}
TAC.df <- calcTAC(fit_maxTHR0.14, assessyear = assess_yr)
THR <- TAC.df[5,2]
TAC <- TAC.df[6,2]/1e3
```

Stock status and trends for SoG Herring are
represented by the ensemble operating model, using the historical time series (1951-2023)
and 6 indicators (Figure \@ref(fig:indicators)).
Ensemble operating model trends are similar to those presented for SoG Herring 
in previous stock assessments (e.g., @dfo2024), with  near-identical data 
sources and choice of indicators.
Step wise comparisons are found in @johnson2024.

(ref:indicators-cap) Estimated stock status indicators for Pacific Herring in the Strait of Georgia (SoG) major stock assessment region (SAR) from `r yr_range[1]` to `r yr_range[2]`, as estimated by the ensemble operating model. All lines indicate weighted ensemble posterior median values, and where applicable, shading indicates `r ci_pct`\% credibility intervals. (A) Catch (x 1,000 t). (B) Estimated spawning stock biomass (x 1,000 t) with limit reference point (LRP; dashed red line), productive period upper stock reference (USR; dashed green line), and unfished spawning biomass ($B_0$; dashed black line). (C) Estimated natural mortality rate ($M$ /yr; red line). (D) Estimated harvest rate ($U_t$, black trend line), with dashed horizontal line at 20\% for 1983-2022. (E) Recruitment of age-`r age_recruit` fish, with historical average shown as a black dashed line. (F) Estimated surplus production (vertical axis) vs. spawning biomass (horizontal axis) for 1988 (triangle) though 2023 (square) with lighter points representing earlier years. Vertical dashed lines (left to right): LRP (red), productive period USR (green), and $B_0$ (black).

```{r indicators, fig.cap = "(ref:indicators-cap)", fig.asp = 1.1}
# Six-panel indicator plot
baseplot_indicators(dat = true_dat)
```

### Historical and Recent Stock Trajectory and Trends {-}

#### Spawning Biomass and Status Relative to Reference Points {-}

```{r OM_stock_status, echo = FALSE, results = "asis"}
stock_status_text(  refPtsTab = ensRefPtsTable,
                    parTab = ensParTable,
                    fYear = 1951, lYear = MSE_yr )
```

The ensemble model estimates spawning biomass to be above $B_{MSY}$ and
all candidate USRs for the majority of the historical time series 
(Figure \@ref(fig:indicators)A).
Additionally, median posterior estimates of spawning biomass have remained
above the LRP in all years since 1970.

#### Recruitment and Natural Mortality {-}

Most of the fluctuations in estimated spawning biomass have been attributed 
to estimated recruitment, with lower than average age-1 recruitment in the 1970s, 
1980s, and in 2007 and 2009, all of which mirror dips in spawning biomass. 
Above average recruitment occurred in the intervening times corresponding to
biomass peaks (Figure \@ref(fig:indicators)E). 
While 2020 is one of the three highest peaks in estimated spawning biomass 
(since 1970), rising estimates of natural mortality (Figure \@ref(fig:indicators)C)
since 2015 has moderated impacts of above average recruitment.
Finally, opposing fluctuations in estimated recruitment and natural mortality 
likely contribute to increased uncertainty in spawning biomass and 
forecast biomass.

#### Spawn Biomass Production {-}

Estimated spawning biomass production was high in 2019 and then declined each 
of the next three year to 2023 (Figure \@ref(fig:indicators)F), despite spawning
biomass remaining fairly constant. This is likely due to the increasing trend
in natural mortality estimates (Figure \@ref(fig:indicators)C).

<!-- #### Fishing Mortality {-} 
JC: not including because not talking about RR-->

### Ecosystem and Climate Change Considerations {-}

Ecosystem considerations are taken into account in the model in a number of ways:

1. Different than most fisheries, the biological LRP is set to a 
higher value than what is recommended by default [@dfo2009] (0.4$B_{MSY}$) and 
0.5$B_{MSY}$ which is implemented in New Zealand fisheries [@shelton2008].
This is due to the middle-of-the-food-chain nature of 
Herring which likely drives observed high variability in Herring productivity.
The higher LRP allows for a certain amount of herring to remain for predators, such as
salmon and whales.
2. While predators, prey and other ecosystem indicators are not directly 
incorporated into models, they are modelled implicitly via time varying natural 
mortality and recruitment. The new density dependent mortality feature
of the new operating model, SISCAH, better represents the underpinning ecosystem interactions.
3. BC Pacific Herring are managed as 5 stock assessment regions such that stock specific 
life history, abundance, and pressures are fed into the models. 
4. Candidate USRs [@dfo2022a], which incorporate area specific ecosystem considerations,
have been introduced to the MSE process and IFMP for consideration as (1) a biomass target
objective and (2) the lower boundary of the Healthy zone.
Ongoing research incorporating spatial, and other ecosystem indicators highlights
DFO's commitment to an ecosystem approach to fisheries management.
Summary of recent research can be found in @dfo2024.

<!--### History of Management Decisions,; Landings/Harvest/Effort, TAC \& Catch Advice] {-}-->

### History of Management {-}

A maximal harvest rate of 20% has been implemented since 1983 for the SoG Herring SAR. 
The realized harvest rate has mostly hovered below this rate since then with reduced harvest
rates applied for the past few years (Figure \@ref(fig:indicators), panel D).
Since 2015 the selection of annual harvest rates has been informed by 
MP simulation testing.

### Projections {-#sec:projections}

#### Interim Year Update for 2024 {-}

```{r current_year_statement, echo = FALSE, warnings = FALSE, results = "asis" }
curr_biomass_text(  mpFit   = fit_maxTHR0.14,
                    fYear   = 1951, 
                    thisYr  = assess_yr,
                    MSEyr   = MSE_yr,
                    B0 = NULL  )
```

#### Application of MP for 2025 {-}

We simulation tested three MPs,
each with hockey-stick shaped harvest control rules
(Figure \@ref(fig:plotHCRules)).
The MP with a maximum target harvest rate of 14% 
meets the conservation objective and
is compliant with both the “DFO Sustainable Fisheries Framework” policy and 
the "Precautionary Approach" (@dfo2009), as per the 
[Terms of Reference](https://www.dfo-mpo.gc.ca/csas-sccs/Schedule-Horraire/2024/09_23-eng.html). 

```{r forecast_statement, echo = FALSE, warnings = FALSE, results = "asis"}
proj_biomass_text(  mpFit = fit_maxTHR0.14,
                    fYear = 1951, 
                    assessYr = assess_yr,
                    B0 = NULL,
                    TAC = TAC,
                    targHR = THR  )
```
This is shown graphically in Figure 4.

<!-- ### Harvest Decision Rule Outputs {-}-->
```{r TACtable, echo = FALSE, warnings = FALSE, eval = FALSE}

TACtabCap <- "Harvest decision rule outputs when the SoG Herring SAR MP is applied to data up to 2024."

kable( TAC.df, escape = FALSE, 
        caption = TACtabCap, booktabs = TRUE,
        align = c("l","c"),
        linesep = "\\addlinespace")
```

```{r MPcaptions, include = FALSE, echo = FALSE}
stockLabs <- fit_maxTHR0.14$stock
species   <- fit_maxTHR0.14$ctlList$ctrl$speciesName
stock     <- fit_maxTHR0.14$ctlList$ctrl$stockNames
```

```{r mpHCR, echo = FALSE, message = FALSE, fig.cap = hcrCaption, fig.asp = 0.67}
hcrCaption <- paste0(
  "Strait of Georgia (SoG)",
  " Pacific Herring management procedure harvest",
  " control rule (line) and total allowable catch (TAC)",
  " advice in ", assess_yr + 1, " $Q_{2025}$",
  " in kilotonnes (kt; annotation).",
  " Point indicates forecast spawning biomass in ", assess_yr + 1,
  " $\\hat{B}_{2025}$ and",
  " harvest rate in ", assess_yr + 1, " $U_{2025}$."
)
plotRule(repList = mpFitList[[MPname]])
#TODO (Sam): pls make this work with MPname 
```

<!--### Evaluation of Exceptional Circumstances {-}-->

### Performance Evaluation for Interim Years and Exceptional Circumstances {-}

Fishery monitoring data collected since the adoption of the current SoG
Herring MP indicate that the population is behaving within operating 
model expectations. There is no indication of operating
model mis-specification (Figure \@ref(fig:ECfig)), also know as exceptional
circumstances. Mis-specification is evaluated visually by comparing catch and 
spawn index values within the central 95% of 2023 ensemble operating 
model simulations.

<!-- For ECfig:
JC: Why is the uncertainty in the index larger than SB?

SDNJ: Because the noise in the index is the uncertainty in SSB (process error) plus 
uncertainty in spawn index (observation error)-->

```{r ECfig, echo = FALSE, fig.cap = ECfigCap, fig.asp = 1.1 }
ECfigCap <- "Evaluation of exceptional circumstances and operating model mis-specification: graphical comparison of simulation envelopes from the ensemble model and catch and spawn index data from interim year(s). Panels show the central 95\\% of projected simulated data, including spawn index data (top) and catch (bottom) under the weighted ensemble 2023 SoG operating model. Forward projections in time (15-years) represent the maxTHR0.14 MP. Realised data are overlaid as points for the history (black) and projection years (red)."

SOGdata <- read.csv("data/SOGIdxCatData.csv")

plotSimDataEnvelopes( obj = mpBlobList[[MPname]],
                      sIdx = 1, pIdx =1, fIdx = 5,
                      Cdata = SOGdata$Catch,
                      Idata = SOGdata$totSpawn,
                      fYear = 1951 )
```
