## ASSESSMENT UNDER MSE {-#sec:assessment}

```{r calcTACchunk, echo = FALSE, include = FALSE}
TAC.df <- calcTAC(fit_maxTHR0.14)

```

Stock trends and status for SoG Herring historical time series (1951-2023)
is represented by the ensemble operating model, 
presented using 6 indicators (Figure \@ref(fig:indicators)).
Ensemble operating model trends are similar to those presented for SoG Herring 
in previous stock assessments (e.g., [@dfo2024]), and data sources are near-identical
as described in @johnson2024.

<!--TODO (Sam): 
For ref:indicators-cap:
1. the autodates need to state 1951-2023. I'm not sure where to change `r yr_range[1]` to `r yr_range[2]` 
2. please confirm age-`r age_recruit` fish is age-1 (I thought this was age-2 but maybe I'm wrong)
3. (F) I think the year labels in the legend are reversed- pls check (we want darkest colour as most recent) -->

(ref:indicators-cap) Estimated stock status indicators for Pacific Herring in the Strait of Georgia (SoG) major stock assessment region (SAR) from `r yr_range[1]` to `r yr_range[2]`, as estimated by the ensemble operating model. All lines indicate weighted ensemble posterior median values, and where applicable, shading indicates `r ci_pct`\% credibility intervals.(A) Catch (black solid line). (B) Estimated spawning stock biomass (x 1,000 t) with limit reference point (LRP; dashed red line), productive period upper stock reference (USR; dashed green line), and unfished spawning biomass ($B_0$; dashed black line). (C) Estimated natural mortality rate ($M$ /yr; red line). (D) Estimated harvest rate ($U_t$, black trend line) with dashed horizontal line at 20\% shown from 1983-2022). (E) Recruitment of age-`r age_recruit` fish. (F) Estimated surplus production (vertical axis) vs. spawning biomass (horizontal axis) with lighter points representing earlier years. vertical dashed lines denote the LRP (red), productive period USR (green), and $B_0$ (black).

```{r indicators, fig.cap = "(ref:indicators-cap)", fig.asp = 1.1}
# Six-panel indicator plot
baseplot_indicators(dat = true_dat)
```


### Historical and Recent Stock Trajectory and Trends {-}

[Mandatory.
Maximum text length one page.
Must be first section in Assessment section.]
[Report trends in relevant indicator(s), one per paragraph,
in the following format:]


#### Spawning Biomass {-}
The ensemble model estimates spawning biomass to be above the USR
for the majority of the historical time series since
Additionally, spawning biomass is estimated to be above the LRP
in all years since 1970.



### History of [Management; Landings/Harvest/Effort, TAC \& Catch Advice] {-}

[Mandatory.
Title can vary.
Maximum text length one page.] 

Harvest options for `r assess_yr+1` reflect application of
simulation-tested MPs for each major SAR,
derived from the Herring OM [@benson2022]. 
OM conditioning was updated in 2022 using
historic stock and fishery data from `r major_start_yr` to 2021;
no MP updates were conducted for 2023.
MPs are not provided for HG because this is now conducted within the 
HG rebuilding plan process.^[Haida Gwaii 'íináang | iinang Pacific Herring:
An ecosystem overview and ecosystem-based rebuilding plan.
Draft consultation period occurred September 26 to December 16, 2022; 
Final approval scheduled for fall 2023.]
<!-- TODO: Need to update rebuilding plan reference -->

<!-- #### History of Hatchery and Supplementation {-} -->

<!-- [If applicable (otherwise, do not include).] -->

<!-- #### History of Freshwater Habitat Impacts {-} -->

<!-- [If applicable (otherwise, do not include).] -->

### Projections {-#sec:projections}

We simulation tested three MPs with hockey-stick shaped harvest control rules (#refSecThisDoc)
and present performance indicators for the MP tuned to meet the conservation objective and
that is compliant with both the “DFO Sustainable Fisheries Framework” policy and 
“A fishery decision-making framework incorporating the Precautionary Approach” policy
(as per the Terms of Reference (#refLinkCSASTOR)).
  

[Mandatory.
Title can vary.
Tables or figures can be used as necessary to support decision-making.]

`r proj_biomass_text()`

<!-- Insert `proj_biomass_text()` function to automate text:

- In the absence of fishing,
spawning biomass in 2024 $\SB_{2024}$ is forecast
to be 73,375 t (posterior median; Table XX).
Spawning biomass in 2024 is forecast to be
below the LRP of $0.3\SB_0$ (41,547 t) with a 8.0% probability,
in the absence of fishing (Table XX and Figure XX).
 -->
<!-- TODO: define three generations == 15-years -->

### Ecosystem and Climate Change Considerations {-}

Density dependence, where increased mortality is associated with smaller biomass
models ecosystem effects such as the increased percentage of natural mortality 
in the event of lower biomass, consistent with predator consumption. 

### Harvest Decision Rule Outputs {-}

```{r, TACtable, echo = FALSE, warnings = FALSE}

TACtabCap <- "Harvest decision rule outputs when the SoG Herring SAR MP is applied to data up to 2024."

kable( TAC.df, escape = FALSE, 
        caption = TACtabCap, booktabs = TRUE,
        align = c("l","c"),
        linesep = "\\addlinespace")


```

```{r, MPcaptions, include = FALSE, echo = FALSE}

stockLabs <- fit_maxTHR0.14$stock
species   <- fit_maxTHR0.14$ctlList$ctrl$speciesName
stock     <- fit_maxTHR0.14$ctlList$ctrl$stockNames

SBtRtMtFtCaption <- paste("Time series of spawning biomass with scaled spawn 
indices (top),recruitments (second row), natural mortality (third row), and
harvest rates (bottom row) for substocks of ", species, ". Stocks are, from 
left to right,", paste(stockLabs, collapse = ", "),".", sep = "")


hcrCaption <- paste0(stock," Herring management procedure harvest control rule 
(line) and TAC advice (annotation).")


```

```{r AtAGlance, include = include_extras, echo = FALSE, message = FALSE, fig.cap = SBtRtMtFtCaption , fig.asp = 1.1}
plotMulti(ts = c("SBtIdx","Rt","Mt","Ftg"), repList = fit_maxTHR0.14)
```

```{r mpHCR, echo = FALSE, message = FALSE, fig.cap = hcrCaption, fig.asp = 1.1}
plotRule(repList = fit_maxTHR0.14)
```

### Evaluation of Exceptional Circumstances/Assessment Triggers {-}

Fishery monitoring data collected since the adoption of the current SoG
Herring MP indicate that the population is behaving within operating 
model expectations. In formal language there is no indication of operating
model mis-specification (Figure \@ref(fig:ECfig)), also know as exceptional
circumstances. Mis-specification is checked visually by comparing catch and 
spawn index values within the central 95% of 2023 ensemble operating 
model simulations.

<!-- TODO (Sam): 
For ECfig:
1. top - Is the grey envelope (post 2023) spawning biomass and the blue spawn index? Why is the
uncertainty in the index larger than SB?
2. Data point labels in both legends are reversed- pls change to:
(top) "Spawning biomass"; Simulated spawn index range"; "Observed index to 2023"; "Observed index post 2023"
(bottom) "Historical and simulated catch"; "Observed catch to 2023"; "Observed catch post 2023"
-->

```{r ECfig, echo = FALSE, fig.cap = ECfigCap, fig.asp = 1.1 }
ECfigCap <- "Evaluation of exceptional circumstances and operating model mis-specification: graphical comparison of simulation envelopes from the ensemble model and catch and spawn index data from interim year(s) f. Panels show the central 95\\% of projected simulated data, including spawn index data (top) and catch (bottom) under the weighted ensemble 2023 SoG operating model. Forward projections in time (15-years) represent the maxTHR0.14 MP. Realised data are overlaid as points for the history (black) and projection years (red)."

SOGdata <- read.csv("data/SOGIdxCatData.csv")

plotSimDataEnvelopes( obj = mpBlobList[[2]],
                      sIdx = 1, pIdx =1, fIdx = 5,
                      Cdata = SOGdata$Catch,
                      Idata = SOGdata$totSpawn,
                      fYear = 1951 )

```
