\clearpage
 

`r if(knitr:::is_latex_output()) '\\Appendices'`
 
`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`
 
# APPENDIX A {-#sec:app-a}

This appendix summarises closed loop simulations evaluating candidate
management procedures for the Strait of Georgia (SOG) herring fishery. SOG herring
is the first BC herring fishery to use the new herring operating model framework, 
named the Spatially Integrated Statistical Catch-at-Age Herring Operating Model 
(SISCAH-OM). Although there are several differences between SISCAH and the 
previous herring assessment and operating models, the main difference in SISCAH-OM
is the ability to model density dependent mortality (DDM), where 
natural mortality is higher when biomass is lower [@johnson2024].

Results from and ensemble of five SISCAH-OMs were weighted to incorporate 
uncertainty in future productivity while evaluating candidate MPs. 
We first summarise the simulation approach,
and then present results with some brief discussion of the management implications.  

# Simulation approach {-}

## Operating models {-}

To take into account uncertainty in productivity at high and low stock sizes 
five SoG Herring operating models where chosen. Productivity
at high stock sizes is influenced by the $M_b$ parameter, which 
represents the average mortality rate at very high biomass (generally above 
unfished levels). At low stock size, productivity is more influenced by the 
stock-recruit relationship's steepness parameter $h$, which is the ratio of 
recruitment at 20\% of unfished to the unfished recruitment $R_0$. Overall $h$ 
is better estimated for stocks where the history includes recovery from very 
low stock sizes. Since this isn't case for SoG, we capture some of the 
uncertainty around $M_b$ and $h$ by considering operating models with 
different values for those parameters. First we used an established 
value of $h=0.70$ [@cleary2018] and used a likelihood profile approach,
estimating operating model parameters across a grid of a $M_b$ values.
A likelihood profile is obtained by plotting negative log likelihood 
values estimated by SISCAH at each grid point, which shows the minimum
value (i.e., the value most in agreement with the data) occurs when
$M_b=0.562$ (Figure \@ref(fig:LPfig)). The same likelihood profile
is used to choose the outer range of $M_b = 0.532$ and $M_b = 0.584$, 
which have equal likelihood values and bound the minimum of $M_b = 0.562$
(Figure \@ref(fig:LPfig)). Finally, two additional operating models
are obtained using the optimal $M_b = 0.562$ value and varying
stock-recruit steepness to a lower level of $h = 0.65$ and an upper
level of $h = 0.75$, based on ranges observed in forage fishes
[@myers2001stock;@hsu2024probable].

```{r LPfig, echo = FALSE, fig.cap = LPfigCap }
LPfigCap <- "Fishery monitoring data likelihood function profiles relative to a range of natural mortality asymptotic lower limit ($M_b$) parameter values. The red point shows the most likely value, while the two vertical dashed lines show two values with equal likelihood chosen to bound the central OM1. All SISCAH model fits here have a steepness value of $h = 0.7$"
filepath <- "data/LP_chooseMb.png"
knitr::include_graphics(filepath)

```

The above process creates a cross design (Table \@ref(tab:parRefPtsTable)) of 
operating models. The so called ensemble of operating models is then combined 
via weighted averaging to give estimates biomass, fishery reference points, and 
current stock status relative to those reference points. The central OM (called OM1, 
$M_b=0.562$ and $h = 0.7$) is given the highest weighting of of 0.34, and the 
remaining weight of 0.66 is equally split among the other operating models 
(OM2 - OM5) giving a weight of 0.165 for the remaining models. Then, model
parameters $\theta$ and derived quantites are found via a weighted average,
e.g.,

\begin{equation}
\theta_{ens} = 0.34 \cdot \theta_{OM_1}+ 0.165 \cdot \sum_{k = 2}^5 \theta_{OM_k},
\end{equation}

where $\theta_{ens}$ is the ensemble weighted average of the parameter values
$\theta_{OM_k}$ from operating model $k = 1,...,5$.

For all operating models, we report leading OM parameters, current biomass,
current stock status relative to unfished biomass, and $MSY$ based reference points (Table \@ref(tab:parRefPtsTable)). The final column also gives the weighted ensemble value of the posterior mean values from each OM.

```{r parRefPtsTable, echo = FALSE, warnings = FALSE}

postParTable <- read.csv("data/omGrid_postPars.csv")
ensParTable <- read.csv("data/ensOM_meanPars.csv") |>
                mutate_if(is.numeric,round,2) |>
                mutate( PBTGtLRP = 0.99 )

parTable <- rbind(postParTable,ensParTable)
parTable$X <- NULL

scenRefPtsTable <- read.csv("data/SOG_allScenRefPts.csv")
ensRefPtsTable <- read.csv("data/SOG_ensRefPts.csv") |>
                  mutate_if(is.numeric, round, 2)

refPtsTable <- rbind(scenRefPtsTable,ensRefPtsTable)
refPtsTable$X <- NULL



allEstTable <- parTable |> 
                left_join(refPtsTable, by = "Scenario") |>
                dplyr::select(  h,
                                Mb,
                                M0,
                                m1,
                                Mbar,
                                qs,
                                qd,
                                qb,
                                R0, 
                                B0,                                 
                                BT,
                                DT = DT.x,
                                PBTGtLRP,
                                Bmsy,
                                Umsy,
                                MSY,
                                Uusr,
                                USR,
                                Yusr,
                                Ucrash) |>
                t()

colnames(allEstTable) <- c("OM 1","OM 2","OM 3","OM 4","OM 5","Ensemble")


rownames(allEstTable) <- c( "$h$",
                            "$M_b$",
                            "$M_0$",
                            "$m_1$",
                            "$\\overline{M}$",
                            "$q_s$",
                            "$q_d$",
                            "$q_{blend}$",
                            "$R_0$",
                            "$B_0$",
                            "$B_{2023}$",
                            "$B_{2023}/B_0$",
                            "$P(B_{2023} > 0.3 B_0)$",
                            "$B_{MSY}$",
                            "$U_{MSY}$",
                            "$MSY$",
                            "$U_{USR}$",
                            "$USR$",
                            "$Y_{USR}$",
                            "$U_{Crash}$")


allEstTabCap <- "SISCAH-OM life-history and management parameter values from the 5 operating models (OMs 1 - 5) for stock-recruit steepness ($h$), asymptotic lower limit on depensatory $M$ ($M_b$, /yr), unfished biomass ($SB_0$, kt), unfished recruitment ($R_0$, 1e6), unfished mortality ($M_0$, /yr), mortality depensation rate ($m_1$), time-averaged mortality ($\\overline{M}$, /yr), surface survey design catchability ($q_s$), spawning biomass in 2023 ($SB_{2023}/SB_0$), stock status ($P(SB_{2023} > 0.3 SB_0)$), spawning biomass at maximum sustainable yield ($SB_{MSY}$, kt), harvest rate targeting maximum sustainable yield ($U_{MSY}$), maximum sustainable yield ($MSY$), upper stock reference ($USR$, kt), harvest rate targeting the upper stock reference ($U_{USR}$), equilibrium yield at the upper stock reference ($Y_{USR}$), harvest rate associated with negative production and higher risk of extirpation ($U_{Crash}$). Uncertainty is shown as the 95\\% credible interval where estimates could be drawn from posterior samples (indicated by two parenthetical values), or half the interquartile range where estimates were drawn from 200 year simulations (one parenthetical value)."

kable( allEstTable, escape = FALSE, 
                      caption = allEstTabCap, booktabs = T,
                      align = rep("r",5),
                      format = kable_format,
                      linesep = "\\addlinespace")  %>%
  kable_styling( latex_options = c("striped", "scale_down","hold_position"),
                 bootstrap_options = c("striped", "hover") )

```

The effect of alternative operating models on the stock-recruitment 
relationship is somewhat subtle. 

```{r SRfig, echo = FALSE, fig.cap = SRmodelCap, fig.asp = 1.1}
SRmodelCap <- "Stock-recruitment relationships and age-1 recruitment for the ensemble of operating models. Relationships are shown for the weighted ensemble model with uncertainty (black curve and grey shaded region) and the posterior median relationships for individual operating models (coloured lines). Annual age-1 recruitments are also shown with uncertainty for the historical 1951 - 2023 period (grey points with vertical segments) and for a single possible future in the 2024 - 2038 projection period (pink points). Finally, dashed lines show ensemble weighted average biomassa and recruitment levels for fishery reference points including unfished biomass (grey), the provisional upper stock reference (blue), and biomass at maximum sustainable yield (MSY, green)."

SRpars <- read.csv("data/omGrid_SRpars.csv")

plotStockRecruitEnvelope( obj = mpBlobList[[2]],
                          SRvals = SRpars,
                          inputBmsy = 39.127,
                          inputUSR = 66.97,
                          inputB0 = 90.23 )
```

## Management Procedures {-}

Management procedures are a combination of data, biomass estimation method (EM), 
and a harvest control rule (HCR). 

All MPs tested here use the same data. The EM that is fit to those data is a 
statistical catch-at-age model with a density independent $M$ (DIM) hypothesis 
[@dfo2023management], which uses a simple random walk to estimate 
time-varying $M$. The DIM hypothesis is similar to the previous model used for 
herring assessments and setting TACs, except that the previous model used
a spline with 16 nodes instead of a simple random walk to estimate 
time-varying $M$. The EM provides annual estimates of unfished biomass and 
a 1-year ahead forecast of spawning biomass.

TACs are derived from EM estimates via the harvest control rule. For
SOG herring, we use a "hockey-stick" shaped function defined
by lower and upper control points, and a maximum target harvest rate
[@cox2013roles]. The SOG herring rules have lower and upper control points 
at 30\% and 60\% of unfished biomass, respectively. When biomass 
is estimated to be below the lower control point the rule sets the harvest
rate to zero. Between control points, the harvest rate is a linear ramp 
from zero to the maximum harvest rate, and above the upper control point the 
rule sets harvest rates to the maximum target harvest rate, called $maxTHR$. 
This report shows results for maximum target harvest rates of 6\% 
and 14\% (Figure \@ref(fig:hcrPlot)).

```{r hcrPlot, echo = FALSE, fig.cap = HCRcap, fig.asp = 0.9}
HCRcap <- "Shape of harvest control rule components of management procedures (MP). The green line represents the MP which meets the conservation objective over the 15-year projection period for the ensemble operating model (with density dependent natural mortality). The red line denotes shape of the previously used 20\\% rule and the purple line shows the DFO default harvest control rule with maximum target harvest rate set at $U_{MSY}$ and control points at $0.4 \\cdot B_{MSY}$ and $0.8 \\cdot B_{MSY}$."

plotHCRules(  LCP_i = c(.3,.3,.172),
              UCP_i = c(.6,.6,.342),
              lowF = .0,
              highF = 1,
              U_i = c(.14,.20,.187) ,
              language = "English")
```

## Performance metrics {-}

First, MPs are quantitatively evaluated against the conservation objective (the primary management objective):

1. $P(B_t > 0.3 B_0) \geq 0.75$, or **avoid the limit reference point (LRP) with high 
probability over three herring generations.**

and next against secondary biomass objective:

2. $P(B_t > USR) \geq 0.5$, or **provisional upper stock reference (USR) with neutral
probability**.

Objective 1, also known as the conservation objective, must be passed by
all Herring management procedures. Objective 2 is a target biomass objective that 
includes the recently described provisional $USR$.

To help fishery managers understand trade-offs among biomass and yield, 
additional quantitative performance metrics are estimated. These metrics 
do not have a minimum or target value like objectives, but give greater 
detail on biomass and yield outcomes of each MP over the
15 year simulation.


1. $P(B_t > B_{MSY})$: The probability that biomass is above $B_{MSY}$.

1. $P(B_t > 0.6 B_0)$: The probability that biomass is above $(0.6 B_0)$

1. $P(B_t > USR)$: The probability that biomass is above the provisional $USR$.

1. $P(U_t > U_{MSY})$: The probability that the effective harvest rate
is above $U_{MSY}$.

1. $\overline{C}$: Median (over replicates) of the average (over years) total
landings.

1. $AAV$: Average annual variation in catch, or the mean percentage difference
in catch from year-to-year.

1. $\overline{B_{t}/B_0}$: Average biomass depletion from 2024 - 2038.

1. ${B_{2038}/B_0}$: Median biomass depletion in 2038.

1. $B_{2038}$: Median biomass in 2038.



Performance metrics are estimated via closed loop feedback simulations
with the following simulation algorithm: 

1. For each Operating Model, initialize a pre-conditioned 
simulation model for the period 1951 – 2023 based on a random
draw from the operating model posterior distribution;

2.  Project the SoG Herring DDM operating model into the future one year 
at a time. For each year in the projection, apply the following:
      i)   Update the time series of commercial catch, catch-at-age, and blended 
      spawn survey data up to time-step t for the stock assessment component of the MP;
      ii)  Use an estimation model (a statistical catch age model with density independent $M$) 
      to produce a 1-year ahead forecast of spawning biomass depletion;
      iii)  Determine the target harvest rate associated with the forecast 
      depletion using the HS 30-60 harvest control rule;
      iv)  Using this target harvest rate calculate the total allowable catch from the 1-year ahead biomass 
      forecast;
      v)  Update the simulated DDM operating model herring population in the with 
      incoming recruitment from the DDM stock-recruit curve with recruitment process errors; 
      density dependent natural mortality; and fishing mortality corresponding 
      to the total allowable catch in the previous step. 
      vi)   Repeat steps 2.i - 2.v until the projection period ends (2038).


3.  Repeat steps 1. and 2. 99 more times;

4.  Calculate quantitative performance statistics across all 
    100 replicates.

# Management between management strategy evalauation years {-}

Once a harvest control rule is chosen through the management strategy evaluation 
process that harvest control rule stays in place until the next round of MSE
occurs. For Pacific herring in B.C. that is 3 years or more. 
In the intervening years an estimate of biomass is obtained via
the less computationally instensive DIM SISCAH model.

# Discussion {-}

A fishery that meets Objective 1 could target a 13\% harvest rate with an 
average yield in the 10 - 11 kt range (Table \@ref(tab:statTable). If management
procedures are also targeting the USR (Objective 2), then harvest rates need
to drop to around 7\% (as predicted by the reference point estimates, Table 
\@ref(tab:parRefPtsTable)), and average catches would be closer to the 6 - 7 kt 
range. Any harvest rate above 13\%, including $U_{MSY}$ estimates for every
OM, do not meet either fishery objective. As such, meeting both fishery objectives
requires accepting that TACs will need to be lower in the future, in contrast
to the average of around 20 kt up to 2019. 

Harvest rates need to be much lower than $U_{MSY}$ to avoid the LRP with high 
probability (75\% or more). This is because the LRP is quite high relative to 
$B_{MSY}$ (Table \@ref(tab:parRefPtsTable), Figure \@ref(fig:simEnv)), 
roughly around 70\% of $B_{MSY \vert OM1}$ for OM1. For comparison, the default
limit reference point in Canadian fisheries policy is 40\% of $B_{MSY}$
or some proxy, although this is largely applied to longer lived groundfish
species with less variable recruitment and lower predation pressures.

The EM based on the Density Independent $M$ model appears to overestimate 
biomass in the 1-year ahead forecast. So called positive assessment errors are 
reflected in the higher than neutral probability the effective harvest rates exceed
the maximum target harvest rate (Table \@ref(tab:statTable), $P(U_t > U_{ref})$,
Figure \@ref(fig:simEnv), bottom row). As such, MPs should aim for slightly 
lower maximum harvest rates than the biomass target implies. For example,
reference points imply a harvest rate of around 8\% (weighted over OMs) to
achieve the USR target in the long term, but positive assessment errors mean
that MPs should target closer to 7\%.

```{r, statTable, echo = FALSE }
#This produces HTML output TODO: FIX to table format for word?
wtdPerfTable <- read.csv("data/wtdTable.csv")

statTable <- wtdPerfTable |>
              filter( mp %in% mps2[2] ) |>
              mutate_if(is.numeric, round, 2) |>
              dplyr::select(  mp,
                              pBtGt.3B0,
                              pBtGtBmsy,
                              pBtGt.4B0,
                              pBtGt.5B0,
                              pBtGt.6B0,
                              pBtGtUSR,
                              pBtGthistAve,
                              pUtGtUmsy,
                              avgCatch_t,
                              catchAAV,
                              aveDep, 
                              DnT,
                              BnT ) |>
              mutate( mp = gsub("maxTHR_", "", mp),
                      avgCatch_t = round(avgCatch_t/1e3,2),
                      pBtGt.3B0 = sapply(X = pBtGt.3B0, FUN = passObj, target = .75),
                      pBtGtUSR  = sapply(X = pBtGtUSR, FUN = passObj, target = .5) )
statTable <- do.call(rbind,list(statTable,statTable,statTable))

statTable[1,"mp"] <- "0.14"
statTable[2,"mp"] <- "Umsy"
statTable[3,"mp"] <- "0.20"

statTable[2,-(1:2)] <- NA
statTable[3,-(1:2)] <- NA

statTable[2:3,2] <- "$X$"


colnames(statTable) <- c( "Max HR",
                          "$0.3 B_0$",
                          "$B_{MSY}$",
                          "$0.4 B_0$",
                          "$0.5 B_0$",
                          "$0.6 B_0$",
                          "$USR$",
                          "$\\overline{B}$",
                          "$U_{MSY}$", # Replace with Umsy
                          "$\\overline{C_t}$",
                          "$AAV$",
                          "$\\overline{B_{t}/B_0}$", 
                          "$B_{2038}/B_0$",
                          "$B_{2038}$"
                          )

statTabCap <- "Performance statistics for MPs differing by maximum target harvest rates indicated in column 1. Objectives that are met by an MP are indicated by a checkmark. $\\overline{B}$ is the average biomass from 1951 to 2023. Subscript $t$ indicates the years 2024 - 2038.  Biomass $B$ and Harvest Rate $U$ metrics (columns 2 -8)  are the probability that $B_t$ or $U_t$ respectively is greater than the value indicated in the header. Average Catch $\\overline{C_t}$ and Final Biomass $B_{2038}$ are in biomass units (kt), and final year $B_{2038}/B_0$ and average projection year $\\overline{B_t/B_0}$ stock status is biomass depletion relative to unfished biomass. For SoG, the USR is calculated as 80% of the average spawning biomass during the productive period, ${0.8B}_{1998:2007}$"


objNames <- c(  " " = 1,
                "Conservation" = 1,
                "Biomass" = 2,
                "Max Removal Reference" = 1,
                "Average Catch" = 1,
                "Catch Variability" = 1,
                "Estimated 2024 Catch" = 1,
                "Stock Status" = 2,
                "Final Biomass" = 1)


kable(  statTable, escape = FALSE, 
        caption = statTabCap, booktabs = T,
        align = c(rep("c",ncol(statTable))),
        format = kable_format,
        linesep = "\\addlinespace")  %>%
  kable_styling( latex_options = c("striped", "scale_down","hold_position"),
                 bootstrap_options = c("striped", "hover") ) |>
  add_header_above( objNames, bold = T ) 

```


```{r, captions, include = FALSE, echo = FALSE}
multiGridTulipCap <- "Simulated time series of projected spawning biomass (top), catch (middle), and harvest rate (bottom) for the MP tuned to the conservation objective (HS30-60_HR 0.14). Median values are shown by the thick black lines, the grey shaded region shows the central 95% of each simulation envelope, and three randomly selected individual replicate traces are shown as thin black lines."

```

```{r simEnv, echo = FALSE, fig.cap = multiGridTulipCap, fig.asp = 1.1}
plotGridTulipBtCtUt(mpBlobList[2],
                    labels = mps2[2],
                    yLimB = c(0,2.5),
                    yLimC = c(0,25),
                    yLimHR = c(0,0.5))
```

