\clearpage
 
`r if(knitr:::is_latex_output()) '\\Appendices'`
 
`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`
 
# APPENDIX A {-#sec:app-a}

This appendix summarises closed loop simulations evaluating candidate
management procedures for the Strait of Georgia (SoG) Herring fishery. SoG Herring
is the first BC Herring fishery to use the new Herring operating model framework, 
named the Spatially Integrated Statistical Catch-at-Age Herring Operating Model 
(SISCAH-OM). Although there are several differences between SISCAH and the 
previous Herring assessment and operating models, the main difference in SISCAH-OM
is the ability to model density dependent mortality (DDM), where 
natural mortality is higher when biomass is lower [@johnson2024].

Results from and ensemble of five SISCAH-OMs were weighted to incorporate 
uncertainty in future productivity while evaluating candidate MPs. 
We first summarise the simulation approach,
and then present results with some brief discussion of the management implications.  

# Simulation approach {-}

## Operating models {-}

To take into account uncertainty in productivity at high and low stock sizes 
five SoG Herring operating models where chosen. Productivity
at high stock sizes is influenced by the $M_b$ parameter, which 
represents the average mortality rate at very high biomass (generally above 
unfished levels). At low stock size, productivity is more influenced by the 
stock-recruit relationship's steepness parameter $h$, which is the ratio of 
recruitment at 20\% of unfished to the unfished recruitment $R_0$. Overall $h$ 
is better estimated for stocks where the history includes recovery from very 
low stock sizes. Since this isn't case for SoG, we capture some of the 
uncertainty around $M_b$ and $h$ by considering operating models with 
different values for those parameters. First we used an established 
value of $h=0.70$ [@cleary2018] and used a likelihood profile approach,
estimating operating model parameters across a grid of a $M_b$ values.
A likelihood profile is obtained by plotting negative log likelihood 
values estimated by SISCAH at each grid point, which shows the minimum
value (i.e., the value most in agreement with the data) occurs when
$M_b=0.562$ (Figure \@ref(fig:LPfig)). The same likelihood profile
is used to choose the outer range of $M_b = 0.532$ and $M_b = 0.584$, 
which have equal likelihood values and bound the minimum of $M_b = 0.562$
(Figure \@ref(fig:LPfig)). Finally, two additional operating models
are obtained using the optimal $M_b = 0.562$ value and varying
stock-recruit steepness to a lower level of $h = 0.65$ and an upper
level of $h = 0.75$, based on ranges observed in forage fishes
[@myers2001stock;@hsu2024probable].

```{r LPfig, echo = FALSE, fig.cap = LPfigCap }
LPfigCap <- "Fishery monitoring data likelihood function profiles relative to a range of natural mortality asymptotic lower limit ($M_b$) parameter values. The red point shows the most likely value, while the two vertical dashed lines show two values with equal likelihood chosen to bound the central OM1. All SISCAH model fits here have a steepness value of $h = 0.7$"
filepath <- "data/LP_chooseMb.png"
knitr::include_graphics(filepath)

```

The above process creates a cross design (Table \@ref(tab:parRefPtsTable)) of 
operating models. The so called ensemble of operating models is then combined 
via weighted averaging to give estimates biomass, fishery reference points, and 
current stock status relative to those reference points. The central OM (called OM1, 
$M_b=0.562$ and $h = 0.7$) is given the highest weighting of of 0.34, and the 
remaining weight of 0.66 is equally split among the other operating models 
(OM2 - OM5) giving a weight of 0.165 for the remaining models. Then, model
parameters $\theta$ and derived quantites are found via a weighted average,
e.g.,

\begin{equation}
\theta_{ens} = 0.34 \cdot \theta_{OM_1}+ 0.165 \cdot \sum_{k = 2}^5 \theta_{OM_k},
\end{equation}

where $\theta_{ens}$ is the ensemble weighted average of the parameter values
$\theta_{OM_k}$ from operating model $k = 1,...,5$.

For all operating models, we report leading OM parameters,
current biomass, current stock status relative to unfished biomass, and
$MSY$ based reference points (Table \@ref(tab:parRefPtsTable)).
The final column also gives the weighted ensemble value
of the posterior mean values from each OM.

```{r parRefPtsTable, echo = FALSE, warnings = FALSE}

postParTable <- read.csv("data/omGrid_postPars.csv")
ensParTable <- read.csv("data/ensOM_meanPars.csv") |>
                mutate_if(is.numeric,round,2) |>
                mutate( PBTGtLRP = ifelse(PBTGtLRP > 0.99, 0.99, .) )

parTable <- rbind(postParTable,ensParTable)
parTable$X <- NULL

scenRefPtsTable <- read.csv("data/SOG_allScenRefPts.csv")
ensRefPtsTable <- read.csv("data/SOG_ensRefPts.csv") |>
                  mutate_if(is.numeric, round, 2)

refPtsTable <- rbind(scenRefPtsTable,ensRefPtsTable)
refPtsTable$X <- NULL



allEstTable <- parTable |> 
                left_join(refPtsTable, by = "Scenario") |>
                dplyr::select(  h,
                                Mb,
                                M0,
                                m1,
                                Mbar,
                                qs,
                                qd,
                                qb,
                                R0, 
                                B0,                                 
                                BT,
                                DT = DT.x,
                                PBTGtLRP,
                                Bmsy,
                                Umsy,
                                MSY,
                                Uusr,
                                USR,
                                Yusr,
                                Ucrash) |>
                t()

colnames(allEstTable) <- c("OM 1","OM 2","OM 3","OM 4","OM 5","Ensemble")

rownames(allEstTable) <- c( "$h$",
                            "$M_b$",
                            "$M_0$",
                            "$m_1$",
                            "$\\overline{M}$",
                            "$q_s$",
                            "$q_d$",
                            "$q_{blend}$",
                            "$R_0$",
                            "$B_0$",
                            "$B_{2023}$",
                            "$B_{2023}/B_0$",
                            "$P(B_{2023} > 0.3 B_0)$",
                            "$B_{MSY}$",
                            "$U_{MSY}$",
                            "$MSY$",
                            "$U_{USR}$",
                            "$USR$",
                            "$Y_{USR}$",
                            "$U_{Crash}$")

allEstTabCap <- "SISCAH-OM life-history and management parameter values from the 5 operating models (OMs 1 - 5) for stock-recruit steepness ($h$), asymptotic lower limit on depensatory $M$ ($M_b$, /yr), unfished biomass ($SB_0$, kt), unfished recruitment ($R_0$, 1e6), unfished mortality ($M_0$, /yr), mortality depensation rate ($m_1$), time-averaged mortality ($\\overline{M}$, /yr), surface survey design catchability ($q_s$), spawning biomass in 2023 ($SB_{2023}/SB_0$), stock status ($P(SB_{2023} > 0.3 SB_0)$), spawning biomass at maximum sustainable yield ($SB_{MSY}$, kt), harvest rate targeting maximum sustainable yield ($U_{MSY}$), maximum sustainable yield ($MSY$), upper stock reference ($USR$, kt), harvest rate targeting the upper stock reference ($U_{USR}$), equilibrium yield at the upper stock reference ($Y_{USR}$), harvest rate associated with negative production and higher risk of extirpation ($U_{Crash}$). Uncertainty is shown as the 95\\% credible interval where estimates could be drawn from posterior samples (indicated by two parenthetical values), or half the interquartile range where estimates were drawn from 200 year simulations (one parenthetical value)."

kable( allEstTable, escape = FALSE, 
                      caption = allEstTabCap, booktabs = T,
                      align = rep("r",5),
                      format = kable_format,
                      linesep = "\\addlinespace")  %>%
  kable_styling( latex_options = c("striped", "scale_down","hold_position"),
                 bootstrap_options = c("striped", "hover") )

```

```{r SRfig, eval = FALSE, echo = FALSE, fig.cap = SRmodelCap, fig.asp = 1.1}
SRmodelCap <- "Stock-recruitment relationships and age-1 recruitment for the ensemble of operating models. Relationships are shown for the weighted ensemble model with uncertainty (black curve and grey shaded region) and the posterior median relationships for individual operating models (coloured lines). Annual age-1 recruitments are also shown with uncertainty for the historical 1951 - 2023 period (grey points with vertical segments) and for a single possible future in the 2024 - 2038 projection period (pink points). Finally, dashed lines show ensemble weighted average biomassa and recruitment levels for fishery reference points including unfished biomass (grey), the provisional upper stock reference (blue), and biomass at maximum sustainable yield (MSY, green)."

SRpars <- read.csv("data/omGrid_SRpars.csv")

plotStockRecruitEnvelope( obj = mpBlobList[[MPname]],
                          SRvals = SRpars,
                          inputBmsy = 39.127,
                          inputUSR = 66.97,
                          inputB0 = 90.23 )

# Figure removed at this time. If the limits on the axis were changed so
# that spawning biomass < 100 kt and recruitment < 2500 then the relationships
# could be seen better

# The effect of alternative operating models on the stock-recruitment
# relationship is somewhat subtle.
```

## Management Procedures {-}

Management procedures are evaluated using three components, data, the 
DDM operating model(OM), and the particular harvest control rule.
Fifteen year simulations are used to produce performance metrics which are
used to evaluate the MP. This procedure occurs only every few years or so. 

For interim years in order to apply a forecast for the following year to the MP, 
a simpler estimation model (EM) using a random walk M (DIM) formulation 
which is blind to the true M~B dynamics is used [@dfo2023management].
The DIM EM is similar to the previous model used for 
Herring assessments and setting TACs prior to 2023, 
except that the previous model used
a spline with 16 nodes instead of a simple random walk to estimate 
time-varying $M$. The EM provides annual estimates of unfished biomass and 
a 1-year ahead forecast of spawning biomass.
Using a DIM EM is appropriate for setting up the MSE as it allows us to evaluate 
population dynamic complexities in the OM, and then use simpler EMs.

TACs are derived from EM estimates via the harvest control rule. For
SoG Herring, we use a "hockey-stick" shaped function defined
by lower and upper control points, and a maximum target harvest rate
[@cox2013roles]. The SoG Herring rules often have lower and upper control points 
at 30\% and 60\% of unfished biomass, respectively. When biomass 
is estimated to be below the lower control point the rule sets the harvest
rate to zero. Between control points, the harvest rate is a linear ramp 
from zero to the maximum harvest rate, and above the upper control point the 
rule sets harvest rates to the maximum target harvest rate, called $maxTHR$. 
This report shows results for MPs with maximum target harvest rates of 
currently recommended 14\%, formerly recommended  20\%, and one 
based on $U_{MSY}$([@dfo2009],Figure \@ref(fig:plotHCRules), Table \@ref(tab:statTable)).

Finally, after the three year cycle (eg, in 2026) the operating model is fully 
updated with new data and comprehensive performance evaluation of MPs is 
again undertaken.

<!--
```{r hcrPlot, echo = FALSE, fig.cap = HCRcap, fig.asp = 0.9}
HCRcap <- "Shape of harvest control rule components of management procedures (MP). The green line represents the MP which meets the conservation objective over the 15-year projection period for the ensemble operating model (with density dependent natural mortality). The red line denotes shape of the previously used 20\\% rule and the purple line shows the DFO default harvest control rule with maximum target harvest rate set at $U_{MSY}$ and control points at $0.4 \\cdot B_{MSY}$ and $0.8 \\cdot B_{MSY}$."

plotHCRules(  LCP_i = c(.3,.3,.172),
              UCP_i = c(.6,.6,.342),
              lowF = .0,
              highF = 1,
              U_i = c(.14,.20,.187) ,
              language = "English")
```
-->
## Performance metrics {-}

First, MPs are quantitatively evaluated against the conservation objective (the primary management objective):

1. $P(B_t > 0.3 B_0) \geq 0.75$, or **avoid the limit reference point (LRP) with high 
probability over three Herring generations.**

Next, MP performance is examined using additional metrics of biomass and yield,
which reflect three additional objectives described in 
Section "[Fishery management objectives](#sec:fish-mgmt-objs)".

To help fishery managers understand trade-offs among biomass and yield, 
additional quantitative performance metrics are estimated.
These metrics do not have a minimum or target value like objectives,
but give greater detail on biomass and yield outcomes of each MP
over the 15-year simulation.

1. $P(B_t > B_{MSY})$: The probability that biomass is above $B_{MSY}$.

1. $P(B_t > 0.6 B_0)$: The probability that biomass is above $(0.6 B_0)$

1. $P(B_t > \overline{B}_{prod})$: The probability that biomass is above $(\overline{B}_{prod})$, also called the provisional $USR$.

1. $P(U_t > U_{MSY})$: The probability that the effective harvest rate
is above $U_{MSY}$.

1. $\overline{C}$: Median (over replicates) of the average (over years) total
landings.

1. $AAV$: Average annual variation in catch, or the mean percentage difference
in catch from year-to-year.

1. $\overline{B_{t}/B_0}$: Average biomass depletion from 2024 - 2038.

1. ${B_{2038}/B_0}$: Median biomass depletion in 2038.

1. $B_{2038}$: Median biomass in 2038.

Performance metrics are estimated via the following closed loop feedback 
simulation algorithm:

1. For each operating model, initialize a pre-conditioned 
simulation model for the period 1951 to 2023 based on a random
draw from the operating model posterior distribution;

2.  Project the SoG Herring DDM operating model into the future one year 
at a time. For each year in the projection, apply the following:
      i)   Update the time series of commercial catch, catch-at-age, and blended 
      spawn survey data up to time-step $t$ for the stock assessment component of the MP;
      ii)  Use an estimation model (a statistical catch age model with density independent $M$) 
      to produce a 1-year ahead forecast of spawning biomass depletion;
      iii)  Determine the target harvest rate associated with the forecast 
      depletion using a harvest control rule;
      iv)  Using this target harvest rate calculate the total allowable catch from the 1-year ahead          biomass forecast;
      v)  Update the simulated DDM operating model Herring population in the with 
      depletion using the harvest control rule;
      vi)  Update the simulated DDM operating model Herring population the with 
      incoming recruitment from the DDM stock-recruit curve with recruitment process errors; 
      density dependent natural mortality; and fishing mortality corresponding 
      to the total allowable catch in the previous step. 
      vii)   Repeat steps 2.i - 2.v until the projection period ends (2038).

3.  Repeat Step 1 and Step 2 99 more times;

4.  Calculate quantitative performance statistics across all 
    100 replicates.

# Discussion {-}

Selecting a MP by tuning to the conservation objective leads to
a harvest control rule recommendation with a 14\% maximum target harvest rate.
If applied annually, the
average yield, using recent 20-year average allocation among gear types,
is 10 - 12 kt (Table \@ref(tab:statTable), Figure \@ref(fig:simEnv)).
Selecting
an MP with a maximum target harvest rate at or below 14\% will achieve the 
conservation objective.

Although additional fishery objectives are presented in 
Section "[Fishery management objectives](#sec:fish-mgmt-objs)", these are not
a full suite of objectives for SoG Herring and are not currently used for
further MP tuning.
If this was undertaken, for example, tuning the MP
to a biomass target of 0.6$B_0$, the maximum target harvest rate and average
annual catch would be lower.

Table \@ref(tab:statTable) includes MPs with maximum target harvest rates
of 20% and $U_{MSY}$, which reflect previously implemented harvest rate and 
the reference removal (RR) defined in @dfo2009.  

However, when tested using an operating model that implements a density 
dependent formulation of time-varying natural mortality, 
neither of these harvest rates meets the conservation
objective when included in a ramped harvest control rule and thus
probabilities and performance statistics do not appear in the 
Table \@ref(tab:statTable).
See Figure \@ref(fig:plotHCRules) for shape of these additional simulation tested
rules.

Simulation-evaluation of MPs show that harvest rates need to be lower 
than $U_{MSY}$ to avoid the LRP with high probability (75\% or more). 
This is because the LRP is high, roughly 70\% of 
$B_{MSY}$ for the ensemble operating model
(Table \@ref(tab:parRefPtsTable), Figure \@ref(fig:simEnv)).  
For comparison, the default LRP in Canadian fisheries policy (@dfo2009) is 
40\% of $B_{MSY}$ or some proxy, although this is largely applied to longer 
lived groundfish species with less variable recruitment and 
lower predation pressures.

The EM implements a density independent $M$ model and appears to overestimate 
biomass in the 1-year ahead forecast.
So called positive assessment errors are 
reflected in the higher than neutral probability the effective harvest rate exceeds
the maximum target harvest rate (Figure \@ref(fig:simEnv), bottom row). 
However the EM is in the simulation-testing process and 
any positive assessment errors are accounted for in the meeting of 
the conservation objective.

```{r statTable, echo = FALSE }
#This produces HTML output TODO: FIX to table format for word?
options(width = 300)
wtdPerfTable <- read.csv("data/wtdTable.csv")

statTable <- wtdPerfTable |>
              filter( mp == MPname ) |>
              mutate_if(is.numeric, round, 2) |>
              dplyr::select(  mp,
                              pBtGt.3B0,
                              pBtGtBmsy,
                              pBtGt.4B0,
                              pBtGt.5B0,
                              pBtGt.6B0,
                              pBtGtUSR,
                              pBtGthistAve,
                              pUtGtUmsy,
                              avgCatch_t,
                              catchAAV,
                              aveDep, 
                              DnT,
                              BnT ) |>
              mutate( mp = gsub("maxTHR_", "", mp),
                      avgCatch_t = round(avgCatch_t/1e3,2),
                      pBtGt.3B0 = sapply(X = pBtGt.3B0, FUN = passObj, target = .75),
                      pBtGtUSR  = sapply(X = pBtGtUSR, FUN = passObj, target = .5) )
statTable <- do.call(rbind,list(statTable,statTable,statTable))

statTable[1,"mp"] <- "0.14"
statTable[2,"mp"] <- "$U_{MSY}$"
statTable[3,"mp"] <- "0.20"

statTable[2,-(1:2)] <- NA
statTable[3,-(1:2)] <- NA

statTable[2:3,2] <- "N"
statTable[statTable == "$\\checkmark$"] <- "Y"

colnames(statTable) <- c( "Max HR",
                          "$0.3 B_0$",
                          "$B_{MSY}$",
                          "$0.4 B_0$",
                          "$0.5 B_0$",
                          "$0.6 B_0$",
                          "$\\overline{B}_{prod}$",
                          "$\\overline{B}$",
                          "$U_{MSY}$", 
                          "$\\overline{C_t}$",
                          "$AAV$",
                          "$\\overline{B_{t}/B_0}$", 
                          "$B_{2038}/B_0$",
                          "$B_{2038}$"
                          )

statTabCap <- "Performance statistics for MPs differing by maximum target harvest rates indicated in column 1. Objectives that are met by an MP are indicated by 'Y', otherwise 'N'. $\\overline{B}$ is the average biomass from 1951 to 2023. Subscript $t$ indicates the years 2024 - 2038.  Biomass $B$ and Harvest Rate $U$ metrics (columns 2 -8)  are the probability that $B_t$ or $U_t$ respectively is greater than the value indicated in the header. Average catch $\\overline{C_t}$ and final biomass $B_{2038}$ are in biomass units (kt), and final year $B_{2038}/B_0$ and average projection year $\\overline{B_t/B_0}$ stock status is biomass depletion relative to unfished biomass. For SoG, $B_{prod}$ is calculated as 80% of the average spawning biomass during the productive period, ${0.8B}_{1998:2007}$. Note: The bottom two rows are included to show these MPs do not meet the conservation objective."

objNames <- c(  " " = 1,
                "Conservation" = 1,
                "Biomass" = 2,
                "Max Removal Reference" = 1,
                "Average Catch" = 1,
                "Catch Variability" = 1,
                "Estimated 2024 Catch" = 1,
                "Stock Status" = 2,
                "Final Biomass" = 1)

kable(  statTable, escape = FALSE, 
        caption = statTabCap, booktabs = T,
        align = c(rep("c",ncol(statTable))),
        format = kable_format,
        linesep = "\\addlinespace")  %>%
  kable_styling( latex_options = c("striped", "scale_down","hold_position"),
                 bootstrap_options = c("striped", "hover") ) |>
  add_header_above( objNames, bold = T ) 
```

```{r captions, include = FALSE, echo = FALSE}
multiGridTulipCap <- "Simulated time series of projected spawning biomass (top), catch (middle), and harvest rate (bottom) for the MP tuned to the conservation objective (HS30-60_HR 0.14). Median values are shown by the thick black lines, the grey shaded region shows the central 95% of each simulation envelope, and three randomly selected individual replicate traces are shown as thin black lines."
```

```{r simEnv, echo = FALSE, fig.cap = multiGridTulipCap, fig.asp = 1.1}
plotGridTulipBtCtUt( mpBlobList[MPname],
                     labels = MPname,
                     yLimB = c(0,2.5),
                     yLimC = c(0,25),
                     yLimHR = c(0,0.5))
```

<!-- TODO (Sam): should the M0 and Mbar be on this figure given this is from EM and DIM?-->

```{r AtAGlance, include = include_extras, echo = FALSE, message = FALSE, fig.cap = SBtRtMtFtCaption, fig.asp = 1.1}
SBtRtMtFtCaption <- paste0(
  "Time series of maximum likelihood estimates of (top to bottom): spawning biomass in kilotonnes (kt)",
  " and depletion (red trend), and blended spawn index,",
  " recruitment in millions of fish (Miss Ages denotes years with no age compoisiton data),",
  " estimated natural mortality (density independent formulation), as well as",
  " harvest rate, and catch (kt)."
)
plotMulti(ts = c("SBtIdx","Rt","Mt","Ftg"), repList = fit_maxTHR0.14)
```
