\clearpage
 

`r if(knitr:::is_latex_output()) '\\Appendices'`
 
`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`
 
# APPENDIX A {#app:a}

this is the operating model description...

# Background

Pacific Herring (*Clupea pallasii*) fisheries in British Columbia (BC), Canada, 
have been managed via a management procedure approach since 2017. Management
procedures are formal, repeatable rules for deriving harvest advice (usually
total allowable catches; TACs) from fishery monitoring data, and are tested
against quantitative fishery management objectives in closed loop simulations.
Simulation results are often tested against multiple uncertainties by
defining multiple simulation scenarios, representing alternative hypotheses
about the state of nature. The process of designing, testing, and choosing a 
management procedure is often called a Management Strategy Evaluation (MSE).

This document summarises closed loop simulations evaluating candidate
management procedures for the Strait of Georgia (SOG) herring fishery. SOG herring
is the first BC herring fishery to use the new herring operating model framework, 
named the Spatially Integrated Statistical Catch-at-Age Herring Operating Model 
(SISCAH-OM). Although there are several differences between SISCAH and the 
previous herring assessment and operating models, the main difference in SISCAH-OM
is the ability to model depensatory density-dependent mortality (DDM), where 
natural mortality 
is higher when biomass is lower. The SISCAH-OM, with the use of appropriate
mortality hypothesis, was accepted as an operating model for all 5 major herring 
stock assessment regions by Fisheries and Oceans, Canada (DFO) via a 
regional peer review review process in 2023.

This document summarises simulation tests for a group of candidate SOG 
management procedures. All candidate MPs are tested against a suite of five 
SISCAH operating models, which incorporate uncertainty in future productivity 
at both low and high stock sizes. We first summarise the simulation approach,
and then present results averaged (with a weighting) over the five operating 
models, with some brief discussion of the management implications.  

# Simulation approach


## Operating models

We define five SOG Herring operating models that test against
uncertainty in productivity at high and low stock sizes. Productivity
at high stock sizes is heavily influenced by the $M_b$ parameter, which 
represents the average mortality rate at very high biomass (generally above 
unfished levels). At low stock size, productivity is more influenced by the 
stock-recruit relationship's steepness parameter $h$, which is the ratio of 
recruitment at 20\% of unfished to the unfished recruitment $R_0$.

We tested three values of each parameter in a cross design (Table 
\@ref(tab:parRefPtsTable)). Steepness $h$ values were taken across a uniform 
grid of 0.65, 0.7, and 0.75, while $M_b$ was chosen via a Likelihood Profile approach 
with a steepness of $h = 0.7$ (Figure \@ref(fig:LPfig)). The grids for each 
individual parameter are intersected at their central value, giving a total of 
five OMs. We are unable to choose a grid of $h$ values via a likelihood profile, 
as there is limited information in SOG data supporting a single 'most likely' 
steepness value. Information on steepness is gathered when stock biomass is low
which has not occurred in SOG's history.

OM weighting favours the central OM (called OM1, $M_b=0.562$ and $h = 0.7$)
with a weight of 0.34. OM1's weight is the average of the weights of each axis in the
cross desgn, with $h$ being uniform (i.e., 0.33) and the $M_b$ value's weight based 
on the relative data likelihood function value (i.e, around 0.35). The remaining
operating models are weighted equally across the remaining  with a weight of 
0.165.

For all operating models, we report leading OM parameters, current biomass and 
stock status, and $MSY$ based reference points (Table \@ref(tab:parRefPtsTable)).

```{r LPfig, echo = FALSE, fig.cap = LPfigCap }
LPfigCap <- "Fishery monitoring data likelihood function profiles relative to a range of natural mortality asymptotic lower limit ($M_b$) parameter values. The red point shows the most likely value, while the two vertical dashed lines show two values with equal likelihood chosen to bound the central OM1. All SISCAH model fits here have a steepness value of $h = 0.7$"
filepath <- "figure/LP_chooseMb.png"
knitr::include_graphics(filepath)

```

## Management Procedures

Management procedures are a combination of data, biomass estimation method (EM), 
and a havest control rule (HCR). 

All MPs tested here use the same data, namely the blended spawn index, commercial
fishery catch, and commercial fishery catch-at-age data. The EM, which is fit to 
those data, is a SISCAH model with a density independent $M$ (DIM) hypothesis, 
which uses a simple random walk to estimate time-varying $M$. The DIM hypothesis 
is similar to the previous model used for herring assessments and setting TACs, 
which used a spline instead of a simple random walk for estimating time-varying
$M$. The EM provides annual estimates of unfished biomass and a 1-year ahead 
forecast of spawning biomass.

TACs are derived from EM estimates via the harvest control rule. For
SOG herring, we use the status quo HS 30-60 rule, a "hockey-stick" shaped 
function with control points at 30\% and 60\% of unfished biomass. When spawning
biomass is estimated to be below the lower control point, the rule sets harvest
rates to zero; when biomass estimates are above the upper control point, the rule
sets harvest rates to the maximum target harvest rate, called $maxTHR$. We show
results for maximum target harvest rates of 6\% and 14\% (Figure \@ref(fig:hcrPlot)).



```{r hcrPlot, echo = FALSE, fig.cap = HCRcap}
HCRcap <- "Two harvest control rules tested for candidate SOG herring MPs, showing maximum target harvest rates $maxTHR$ of 6\\% and 14\\%."

plotHCRex(maxSeq = c(0.06,0.14))
```


## Performance metrics

MPs are evaluated against two quantitative management objectives, listed below.

1. $P(B_t > 0.3 B_0) \geq 0.75$, or **avoid the limit reference point (LRP) with high 
probability over three herring generations.**

2. $P(B_t > USR) \geq 0.5$, or **target the upper stock reference (USR) with neutral
probability**.

Objective 1, also known as the conservation objective, must be passed by
all herring management procedures. Objective 2 is a target objective related 
to a recently described provisional $USR$.

To help fishery managers understand trade-offs among biomass and yield, 
additional quantitative performance metrics are estimated. These metrics 
do not have a minimum or target value like objectives, but give greater 
detail on biomass and yield outcomes of each MP over the
15 year simulation.


1. $P(B_t > B_{MSY})$: The probability that biomass is above $B_{MSY}$.

2. $P(U_t > U_{MSY})$: The probability that the effective harvest rate
is above $U_{MSY}$. 

3. $P(C_t > 650 t)$: The probability of a viable fishery where catch is
greater than 650 tonnes.

4. $C_{5pc}$: Catch risk, measured as the fifth percentile of catch over all years and replicates.

5. $\overline{C}$: Median (over replicates) of the average (over years) total
landings.

6. $AAV$: Average annual variation in catch, or the mean percentage difference
in catch from year-to-year.

7. $C_{2024}$: The MP's TAC in 2024.

8. $\overline{B_{t}/B_0}$: Average biomass depletion from 2024 - 2038.

9. ${B_{2038}/B_0}$: Median biomass depletion in 2038.

10. $B_{2038}$: Median biomass in 2038.


<!-- Add average depletion -->
Performance metrics are estimated via closed loop feedback simulations
with the following simulation algorithm: 

1. For each Operating Model, initialize a pre-conditioned 
simulation model for the period 1951 – 2023 based on a random
draw from the operating model posterior distribution;

2.  Project the operating model population and fishery 
into the future one time-step at a time. At each step, 
apply the following:
      i)   Generate the catch and spawn survey data available 
      for stock assessment;
      ii)  Determine 1-year ahead forecast of spawning biomass depletion 
      level from a Density Independent $M$ SISCAH estimation model;
      iii) Apply the HS 30-60 harvest control rule wuth to generate a
      catch limit according to the maximum target harvest rate;
      iv)  Update the operating model population given the catch 
           limits for each fishery and new recruitment;
      v)   Repeat steps 2.i - 2.iv until the projection period ends (2038).

3.  Repeat steps 1. and 2. 99 more times;

4.  Calculate quantitative performance statistics above across all 
    100 replicates.


# MP performance and discussion

A fishery that meets Objective 1 could target a 13\% harvest rate with an 
average yield in the 10 - 11 kt range (Table \@ref(tab:statTable). If management
procedures are also targeting the USR (Objective 2), then harvest rates need
to drop to around 7\% (as predicted by the reference point estimates, Table 
\@ref(tab:parRefPtsTable)), and average catches would be closer to the 6 - 7 kt 
range. Any harvest rate above 13\%, including $U_{MSY}$ estimates for every
OM, do not meet either fishery objective. As such, meeting both fishery objectives
requires accepting that TACs will need to be lower in the future, in contrast
to the average of around 20 kt up to 2019. 

Harvest rates need to be much lower than $U_{MSY}$ to avoid the LRP with high 
probability (75\% or more). This is because the LRP is quite high relative to 
$B_{MSY}$ (Table \@ref(tab:parRefPtsTable), Figure \@ref(fig:simEnv)), 
roughly around 70\% of $B_{MSY \vert OM1}$ for OM1. For comparison, the default
limit reference point in Canadian fisheries policy is 40\% of $B_{MSY}$
or some proxy, although this is largely applied to longer lived groundfish
species with less variable recruitment and lower predation pressures.

The EM based on the Density Independent $M$ model appears to overestimate 
biomass in the 1-year ahead forecast. So called positive assessment errors are 
reflected in the higher than neutral probability the effective harvest rates exceed
the maximum target harvest rate (Table \@ref(tab:statTable), $P(U_t > U_{ref})$,
Figure \@ref(fig:simEnv), bottom row). As such, MPs should aim for slightly 
lower maximum harvest rates than the biomass target implies. For example,
reference points imply a harvest rate of around 8\% (weighted over OMs) to
achieve the USR target in the long term, but positive assessment errors mean
that MPs should target closer to 7\%.

The metrics for the probability of a viable fishery and the 5th percentile of 
simulated TACs are closely related (Table \@ref(tab:statTable), $C_t > 650 t$
and $C_{5pc}$). Each shows that increasing harvest rates generally means accepting 
a higher proportion of years with low catch. Additionally, both metrics appear to 
show an optimal range of 4\% - 6\% harvest rates where the risk of poor fishing 
is lowest. The $C_{5pc}$ column is more informative, however, as it uses absolute 
landings estimates that have a broader range than the more intangible probability 
of viable fishing.




